---
title: "Relay Cooperation: Reproducible Analysis"
author: "A. Visine"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
---

# Setup

This script reproduces the principal table and figures in manuscript order:
Table 1, Figure 4, Figure 5, Figure 6, followed by supplementary analyses.

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

# Load data

```{r load-data}
Waypoints_relay_coop <- read_delim(
  "Waypoints_relay_coop.csv",
  col_names = TRUE,
  delim = ","
)

focal_follow_relay_coop <- read_delim(
  "focal_follow_relay_coop.csv",
  col_names = TRUE,
  delim = ","
)

infant_caregiver <- read_delim(
  "infant_caregiver.csv",
  col_names = TRUE,
  delim = ","
)

basket_carrier <- read_delim(
  "basket_carrier.csv",
  col_names = TRUE,
  delim = ","
)
```

# Transformation

## Trips with girls

```{r}

#List of trips where girls are present (13 trips)

trip_with_girls <- which(focal_follow_relay_coop$n_girls > 0)

# Explicit trip IDs for analyses that subset by trip_id
trips_with_girls <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_girls > 0
]

```

\

## Bouts tables

These tables we be used to produce figure 5.

\

### Resting bouts

```{r}

rest_bout <- data.frame(
    trip_id   = c("TEST"),
    rest = c("TEST"),
    duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  for (step_rest in 2:nrow(nursing_w)-1) {
    
    rest_bout <- rbind(rest_bout, c(woman,
                                        nursing_w$resting_behaviour[step_rest],
                                        nursing_w$Time[step_rest+1] - nursing_w$Time[step_rest]))
  
    
  }
}

#Delete TEST row
rest_bout <- rest_bout[-1,]

#Make sure it is numerical now
rest_bout$duration <- as.numeric(rest_bout$duration)

#Make sure it is logical now
#rest_bout$rest <- as.logical.factor(rest_bout$duration)


```

\

### Foraging bouts

- Foraging

```{r}

foraging_bout <- data.frame(
    trip_id   = c("TEST"),
    forages = c("TEST"),
    duration = c("TEST"))

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  #Keep one woman after the other
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == nursing_woman)
  
  #Keep the first and last row, to know when the trip is ending, for duration calculation
  first_row <- nursing_w[1,]
  last_row <- nursing_w[nrow(nursing_w),]
  
  #Keep only changes in forages
  nursing_w <- subset(nursing_w, is.na(nursing_w$foraging) == FALSE)
  
  #Add back the last row
  nursing_w <- rbind(first_row, nursing_w, last_row)
  
  #First bout where no foraging
  foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[2] - nursing_w$Time[1]))
  
  #For each change in foraging
  for (change in 4:nrow(nursing_w)-2) {
    #If the mother forages, add TRUE in the bout file
    if(nursing_w$foraging[change] == "start"){
      foraging_bout <- rbind(foraging_bout, c(nursing_woman, TRUE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    } else { #if there is no forages, add their name, we'll replace by boy/girl/etc later
foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[change+1] - nursing_w$Time[change]))
}
}

#Last bout where no foraging
foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[nrow(nursing_w)] - nursing_w$Time[nrow(nursing_w)-1]))

}  


#Delete the TEST row
foraging_bout <- foraging_bout[-1,]

#Make sure it is numerical now
foraging_bout$duration <- as.numeric(foraging_bout$duration)


```

\

### Walking bouts

- Waking

```{r}

walking_bout <- data.frame(
  trip_id   = c("TEST"),
  walks = c("TEST"),
  duration = c("TEST"))

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  #Keep one woman after the other
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == nursing_woman)
  
  
  #Keep only changes in walking
  nursing_w <- subset(nursing_w, is.na(nursing_w$walking) == FALSE)
  
  
  #For each change in walking
  for (change in 2:nrow(nursing_w)-1) {
    #If the mother walks, add TRUE in the bout file
    if(nursing_w$walking[change] == "start"){
      walking_bout <- rbind(walking_bout, c(nursing_woman, TRUE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    } else { #if there is no walks, add their name, we'll replace by boy/girl/etc later
      walking_bout <- rbind(walking_bout, c(nursing_woman, FALSE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    }
  }
  
}  

#Delete the TEST row
walking_bout <- walking_bout[-1,]

#Make sure it is numerical now
walking_bout$duration <- as.numeric(walking_bout$duration)


```

\

### Labour bouts

```{r}

labour_bout <- data.frame(
  trip_id   = c("TEST"),
  labour = c("TEST"),
  duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  for (step_labour in 2:nrow(nursing_w)-1) {
    
    labour_bout <- rbind(labour_bout, c(woman,
                                        nursing_w$labour[step_labour],
                                        nursing_w$Time[step_labour+1] - nursing_w$Time[step_labour]))
  }
}


#Delete the TEST row
labour_bout <- labour_bout[-1,]

#Make sure it is numerical now
labour_bout$duration <- as.numeric(labour_bout$duration)


```

\

### Breastfeeding

```{r}

#Although this is coded as "bouts" as it was before, the duration obtained corresponds to the START of breastfeeding behaviour from the beginning of the trip

breastfeeding_bouts <- data.frame(
  trip_id   = c("TEST"),
  breastfeed = c("TEST"),
  duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  time_initial <- nursing_w$Time[1]
  
  for (step_breast in 2:nrow(nursing_w)-1) {
    
    if(nursing_w$breastfeed[step_breast] == TRUE){
      
      breastfeeding_bouts <- rbind(breastfeeding_bouts, c(woman,
                                                          TRUE,
                                                          nursing_w$Time[step_breast] - time_initial))
    }
  }
}

breastfeeding_bouts <- breastfeeding_bouts[-1,]

#Make sure it is numerical now
breastfeeding_bouts$duration <- as.numeric(breastfeeding_bouts$duration)



```


### Verification

```{r}

#Test, all ok

stopifnot(sum(walking_bout$duration, na.rm = TRUE) == sum(labour_bout$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(infant_caregiver$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(foraging_bout$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(rest_bout$duration, na.rm = TRUE))

```

# Table 1: Assistance during foraging trips

This analysis calculates the frequency and duration of infant-care and
basket-carrying assistance during focal women's foraging trips.

For age- and gender-specific categories, calculations include only trips in
which at least one member of that category was present. A category that was
present but did not provide assistance contributes a value of zero. Trips in
which the category was absent are excluded.

For the pooled ``Any other group member'' category, calculations are restricted
to the 17 trips in which the focal woman was accompanied by at least one other
group member.


## 1. Prepare infant-care data

The unknown infant-care category (`?`) is combined with `Ground`.

```{r prepare-infant-data}
infant_caregiver2 <- infant_caregiver %>%
  mutate(
    caregiver = ifelse(caregiver == "?", "Ground", caregiver)
  )
```

## 2. Identify trips in which each category was present

```{r category-presence}
trips_women_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_women > 0
]

trips_men_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_men > 0
]

trips_girls_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_girls > 0
]

trips_boys_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_boys > 0
]

trips_accompanied <- focal_follow_relay_coop %>%
  filter(
    n_women > 0 |
    n_men > 0 |
    n_girls > 0 |
    n_boys > 0
  ) %>%
  pull(ID_focal)
```

### Check denominators

These should reproduce the denominators used in Table 1.

```{r check-denominators}
length(trips_accompanied)    # expected: 17
length(trips_women_present)  # expected: 14
length(trips_men_present)    # expected: 6
length(trips_girls_present)  # expected: 13
length(trips_boys_present)   # expected: 11
```

# Infant care

## 3. Calculate trip-level infant-care percentages

For each trip, calculate the percentage of total trip duration associated with
each infant-care category.

```{r infant-trip-percentages}
caregiver_trip <- infant_caregiver2 %>%
  group_by(trip_id, caregiver) %>%
  summarise(
    duration = sum(duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(trip_id) %>%
  mutate(
    percent = duration / sum(duration, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```

## 4. Infant care by category, conditional on presence

If a category was present but did not provide infant care, the trip contributes
0%. If the category was absent, the trip is excluded.

```{r infant-summary-function}
summarise_care_when_present <- function(category, trips_present) {

  x <- data.frame(trip_id = trips_present) %>%
    left_join(
      caregiver_trip %>%
        filter(caregiver == category) %>%
        select(trip_id, percent),
      by = "trip_id"
    ) %>%
    mutate(
      percent = replace_na(percent, 0)
    )

  x %>%
    summarise(
      caregiver = category,
      N_present = n(),
      N_helped = sum(percent > 0),
      pct_trips_helped = N_helped / N_present * 100,
      mean = mean(percent),
      sd = sd(percent),
      min = min(percent),
      max = max(percent)
    )
}
```

```{r infant-category-summary}
caregiver_summary_present <- bind_rows(
  summarise_care_when_present("woman",  trips_women_present),
  summarise_care_when_present("father", trips_men_present),
  summarise_care_when_present("girl",   trips_girls_present),
  summarise_care_when_present("boy",    trips_boys_present)
)

caregiver_summary_present
```

Expected frequencies:

- Other women: 8/14
- Fathers: 4/6
- Girls: 9/13
- Boys: 4/11

## 5. Infant-care assistance from any other group member

`mother` and `Ground` are excluded when defining assistance from another
individual.

```{r infant-any-helper}
infant_help_trip <- infant_caregiver2 %>%
  filter(trip_id %in% trips_accompanied) %>%
  mutate(
    other_help = !caregiver %in% c("mother", "Ground")
  ) %>%
  group_by(trip_id) %>%
  summarise(
    total_duration = sum(duration, na.rm = TRUE),
    other_duration = sum(duration[other_help], na.rm = TRUE),
    percent_other = other_duration / total_duration * 100,
    .groups = "drop"
  )

infant_help_summary <- infant_help_trip %>%
  summarise(
    N = n(),
    N_helped = sum(other_duration > 0),
    percent_trips_helped = N_helped / N * 100,
    mean = mean(percent_other),
    sd = sd(percent_other),
    min = min(percent_other),
    max = max(percent_other)
  )

infant_help_summary
```

Expected result: **17/17 (100%)**, mean = **38.9%**, SD = **27.0%**,
range = **1.9--100.0%**.

# Basket carrying

## 6. Calculate trip-level basket-carrying percentages

```{r basket-trip-percentages}
basket_trip <- basket_carrier %>%
  group_by(trip_id, caregiver) %>%
  summarise(
    duration = sum(duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(trip_id) %>%
  mutate(
    percent = duration / sum(duration, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```

## 7. Basket carrying by category, conditional on presence

```{r basket-summary-function}
summarise_basket_when_present <- function(category, trips_present) {

  x <- data.frame(trip_id = trips_present) %>%
    left_join(
      basket_trip %>%
        filter(caregiver == category) %>%
        select(trip_id, percent),
      by = "trip_id"
    ) %>%
    mutate(
      percent = replace_na(percent, 0)
    )

  x %>%
    summarise(
      caregiver = category,
      N_present = n(),
      N_carried = sum(percent > 0),
      pct_trips_carried = N_carried / N_present * 100,
      mean = mean(percent),
      sd = sd(percent),
      min = min(percent),
      max = max(percent)
    )
}
```

```{r basket-category-summary}
basket_summary_present <- bind_rows(
  summarise_basket_when_present("woman",  trips_women_present),
  summarise_basket_when_present("father", trips_men_present),
  summarise_basket_when_present("girl",   trips_girls_present),
  summarise_basket_when_present("boy",    trips_boys_present)
)

basket_summary_present
```

Expected frequencies:

- Other women: 4/14
- Fathers: 0/6
- Girls: 8/13
- Boys: 3/11

## 8. Basket-carrying assistance from any other group member

```{r basket-any-helper}
basket_help_trip <- basket_carrier %>%
  filter(trip_id %in% trips_accompanied) %>%
  mutate(
    other_help = !caregiver %in% c("mother", "Ground")
  ) %>%
  group_by(trip_id) %>%
  summarise(
    total_duration = sum(duration, na.rm = TRUE),
    other_duration = sum(duration[other_help], na.rm = TRUE),
    percent_other = other_duration / total_duration * 100,
    .groups = "drop"
  )

basket_help_summary <- basket_help_trip %>%
  summarise(
    N = n(),
    N_helped = sum(other_duration > 0),
    percent_trips_helped = N_helped / N * 100,
    mean = mean(percent_other),
    sd = sd(percent_other),
    min = min(percent_other),
    max = max(percent_other)
  )

basket_help_summary
```

Expected result: **12/17 (70.6%)**, mean = **18.5%**, SD = **21.2%**,
range = **0.0--67.7%**.

# Final checks

The frequencies calculated from the duration data should reproduce the
frequencies reported in Table 1.

```{r final-checks}
caregiver_summary_present %>%
  select(caregiver, N_present, N_helped, pct_trips_helped)

basket_summary_present %>%
  select(caregiver, N_present, N_carried, pct_trips_carried)

infant_help_summary

basket_help_summary
```

# Figure 4

\

### Childcare duration

```{r}

childcare_duration <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  infant <- subset(infant_caregiver,
                   trip_id == nursing_woman)
  
  mother_duration <- sum(infant$duration[infant$caregiver == "mother"], na.rm = TRUE)
  ground_duration  <- sum(infant$duration[infant$caregiver == "Ground"], na.rm = TRUE)
  other_duration  <- sum(infant$duration[!infant$caregiver %in% c("mother", "Ground")], na.rm = TRUE)
  
  childcare_duration <- rbind(
    childcare_duration,
    data.frame(
      trip_id         = nursing_woman,
      mother_duration = mother_duration,
      ground_duration = ground_duration,
      other_duration  = other_duration
    )
  )
}

childcare_duration
childcare_duration[trip_with_girls,]

```

```{r}

childcare_duration_detailed <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  infant <- subset(infant_caregiver,
                   trip_id == nursing_woman)
  
  by_caregiver <- aggregate(duration ~ caregiver, data = infant, sum)
  by_caregiver$trip_id <- nursing_woman
  
  childcare_duration_detailed <- rbind(childcare_duration_detailed, by_caregiver)
}

childcare_duration_detailed

```

```{r}

caregiver_colours <- c(
  "mother"  = "#85abd2",
  "woman"   = "#85d2b9",
  "girl"    = "#85d270",
  "boy"     = "#d29485",
  "father"  = "#d294b3",
  "ground"  = "#d2ab85",
  "unknown" = "grey"
)

childcare_frac <- childcare_duration_detailed %>%
  group_by(trip_id) %>%
  mutate(frac = duration / sum(duration)) %>%
  ungroup()

# Rename ? to unknown
childcare_frac$caregiver[childcare_frac$caregiver %in% c("Ground", "?")] <- "ground"

# Set factor order (bottom to top in stacked bar)
childcare_frac$caregiver <- factor(
  childcare_frac$caregiver,
  levels = c("ground", "father", "boy", "girl", "woman", "mother")
)

# Assign trip numbers 1 to 18 in the order they appear
childcare_frac$trip_number <- as.numeric(factor(childcare_frac$trip_id,
                                                levels = unique(childcare_frac$trip_id)))
childcare_frac$trip_number <- factor(childcare_frac$trip_number)

ggplot(childcare_frac,
       aes(x = trip_number, y = frac, fill = caregiver)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = caregiver_colours) +
  labs(
    x = "Focal trip",
    y = "Proportion of trip duration",
    fill = "Category",
    title = "A. Infant care"
  ) +
  theme_minimal() +
  theme(# Plot title
    plot.title = element_text(size = 16),
    
    # Axis titles
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    
    # Axis tick labels (numbers/text)
    axis.text.x = element_text(size = 16, angle = 0, hjust = 1),
    axis.text.y = element_text(size = 16),
    
    # Legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

ggsave("duration_prop_childcare.pdf",
       width = 10, height = 6, dpi = 300)

```

\

```{r}

#Just mother percentage
summary(c(childcare_frac$frac[which(childcare_frac$caregiver == "mother")], 0))
sd(c(childcare_frac$frac[which(childcare_frac$caregiver == "mother")], 0))

```

\

### Basket duration

```{r}

basket_duration_detailed <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  basket <- subset(basket_carrier,
                   trip_id == nursing_woman)
  
  by_caregiver <- aggregate(duration ~ caregiver, data = basket, sum)
  by_caregiver$trip_id <- nursing_woman
  
  basket_duration_detailed <- rbind(basket_duration_detailed, by_caregiver)
}

basket_duration_detailed

```

```{r}

basket_frac <- basket_duration_detailed %>%
  group_by(trip_id) %>%
  mutate(frac = duration / sum(duration)) %>%
  ungroup()

# Rename to match colour vector
basket_frac$caregiver[basket_frac$caregiver %in% c("Ground", "?")] <- "ground"

# Set factor order (bottom to top in stacked bar)
basket_frac$caregiver <- factor(
  basket_frac$caregiver,
  levels = c("ground", "father", "boy", "girl", "woman", "mother")
)

# Assign trip numbers 1 to 18 in the order they appear
basket_frac$trip_number <- as.numeric(factor(basket_frac$trip_id,
                                             levels = unique(basket_frac$trip_id)))
basket_frac$trip_number <- factor(basket_frac$trip_number)

ggplot(basket_frac,
       aes(x = trip_number, y = frac, fill = caregiver)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = caregiver_colours) +
  labs(
    x = "Focal trip",
    y = "Proportion of trip duration",
    fill = "Category",
    title = "B. Basket carrying",
    size = 16
  ) +
  theme_minimal() +
  theme(# Plot title
    plot.title = element_text(size = 16),
    
    # Axis titles
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    
    # Axis tick labels (numbers/text)
    axis.text.x = element_text(size = 16, angle = 0, hjust = 1),
    axis.text.y = element_text(size = 16),
    
    # Legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

ggsave("duration_prop_basket.pdf",
       width = 10, height = 6, dpi = 300)

```

\

```{r}

#Just mother percentage
summary(basket_frac$frac[which(basket_frac$caregiver == "mother")])
sd(basket_frac$frac[which(basket_frac$caregiver == "mother")])

#Just ground percentage
summary(basket_frac$frac[which(basket_frac$caregiver == "ground")])
sd(basket_frac$frac[which(basket_frac$caregiver == "ground")])

```

\

### On the ground

```{r}

basket_floor_pct <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  basket <- subset(basket_carrier,
                   trip_id == nursing_woman)
  
  total_duration <- sum(basket$duration, na.rm = TRUE)
  floor_duration <- sum(basket$duration[basket$caregiver == "Ground"], na.rm = TRUE)
  
  basket_floor_pct <- rbind(
    basket_floor_pct,
    data.frame(
      trip_id    = nursing_woman,
      floor_duration = floor_duration,
      total_duration = total_duration,
      floor_pct  = floor_duration / total_duration * 100
    )
  )
}

basket_floor_pct
summary(basket_floor_pct$floor_pct)
sd(basket_floor_pct$floor_pct)

```

\

# Figure 5

The plotting code below generates the temporal caregiving/activity panels. The following subsections reproduce the childcare- and basket-transfer statistics reported in the Results.


The plots:
  
  ```{r}

caregiver_colors <- c(
  # "boy"    = "#d29485",
  "girl"   = "#85d285",
  #"woman"  = "#85d2b3",
  "other"  = "grey",
  "mother" = "#85abd2",
  "Ground" = "#d2ab85"
)

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  subtable_info <- subset(infant_caregiver, infant_caregiver$trip_id == nursing_woman)
  subtable_info_basket <- subset(basket_carrier, basket_carrier$trip_id == nursing_woman)
  subtable_info_forage <- subset(foraging_bout, foraging_bout$trip_id == nursing_woman)
  subtable_walking <- subset(walking_bout, trip_id == nursing_woman)
  subtable_labour <-  subset(labour_bout, trip_id == nursing_woman)
  subtable_rest <-  subset(rest_bout, trip_id == nursing_woman)
  subtable_breastfeed <- subset(breastfeeding_bouts, trip_id == nursing_woman & breastfeed == 1)
  
  
  # Compute cumulative positions
  subtable_info$frac <- subtable_info$duration / sum(subtable_info$duration)
  subtable_info$xmax <- cumsum(subtable_info$frac)
  subtable_info$xmin <- c(0, head(subtable_info$xmax, -1))
  
  subtable_info_basket$frac <- subtable_info_basket$duration / sum(subtable_info_basket$duration)
  subtable_info_basket$xmax <- cumsum(subtable_info_basket$frac)
  subtable_info_basket$xmin <- c(0, head(subtable_info_basket$xmax, -1))
  
  subtable_info_forage$frac <- subtable_info_forage$duration / sum(subtable_info_forage$duration)
  subtable_info_forage$xmax <- cumsum(subtable_info_forage$frac)
  subtable_info_forage$xmin <- c(0, head(subtable_info_forage$xmax, -1))
  
  subtable_walking$frac <- subtable_walking$duration / sum(subtable_walking$duration)
  subtable_walking$xmax <- cumsum(subtable_walking$frac)
  subtable_walking$xmin <- c(0, head(subtable_walking$xmax, -1))
  
  subtable_labour$frac <- subtable_labour$duration / sum(subtable_labour$duration)
  subtable_labour$xmax <- cumsum(subtable_labour$frac)
  subtable_labour$xmin <- c(0, head(subtable_labour$xmax, -1))
  
  subtable_rest$frac <- subtable_rest$duration / sum(subtable_rest$duration)
  subtable_rest$xmax <- cumsum(subtable_rest$frac)
  subtable_rest$xmin <- c(0, head(subtable_rest$xmax, -1))
  
  subtable_breastfeed$x <- subtable_breastfeed$duration / sum(subtable_info$duration)
  
  # Keep only walking and foraging TRUE
  subtable_walking <- subset(subtable_walking, walks == 1)
  subtable_info_forage <- subset(subtable_info_forage, subtable_info_forage$forages == 1)
  subtable_labour <- subset(subtable_labour, labour == 1)
  subtable_rest <- subset(subtable_rest, rest == 1)
  
  #Keep only mother and girls
  #subtable_info <- subset(subtable_info, subtable_info$caregiver %in% c("mother", "girl", "Ground"))
  #subtable_info_basket <- subset(subtable_info_basket, subtable_info_basket$caregiver %in% c("mother", "girl", "Ground"))
  
  #Assign bar position
  bar_height <- 0.05
  
  y_infant   <-  0.20
  y_forage   <-  0.10
  y_basket   <-  0.00
  
  
  subtable_info$ymin <- y_infant - bar_height
  subtable_info$ymax <- y_infant + bar_height
  
  subtable_info_forage$ymin <- y_forage - bar_height
  subtable_info_forage$ymax <- y_forage + bar_height
  
  subtable_walking$ymin <- y_forage - bar_height
  subtable_walking$ymax <- y_forage + bar_height
  
  subtable_labour$ymin <- numeric(nrow(subtable_labour))
  subtable_labour$ymax <- numeric(nrow(subtable_labour))
  
  if (nrow(subtable_labour) > 0) {
    subtable_labour$ymin <- y_forage - bar_height
    subtable_labour$ymax <- y_forage + bar_height
  }
  
  subtable_rest$ymin <- numeric(nrow(subtable_rest))
  subtable_rest$ymax <- numeric(nrow(subtable_rest))
  
  if (nrow(subtable_rest) > 0) {
    subtable_rest$ymin <- y_forage - bar_height
    subtable_rest$ymax <- y_forage + bar_height
  }
  
  subtable_info_basket$ymin <- y_basket - bar_height
  subtable_info_basket$ymax <- y_basket + bar_height
  
  
  # Plot
  p <- ggplot() +
    
    # Infant holding
    geom_rect(
      data = subtable_info,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax,
          fill = caregiver)
    ) +
    
    geom_segment(
      data = subtable_breastfeed,
      aes(
        x = x, xend = x,
        y = y_infant - bar_height,
        yend = y_infant + bar_height
      ),
      colour = "darkblue",
      linetype = "dotted",
      linewidth = 0.8
    ) +
    
    # Walking (overlay on same bar)
    geom_rect(
      data = subtable_walking,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#ffd690",
      alpha = 1
    ) +
    
    # Rest
    geom_rect(
      data = subtable_rest,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#b8dcb8",
      alpha = 1
    ) +
    
    # Labour (background bar)
    geom_rect(
      data = subtable_labour,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#c3b0d5",
      alpha = 1
    ) +
    
    # Foraging
    geom_rect(
      data = subtable_info_forage,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#756a80",
      alpha = 1
    ) +
    
    # Basket carrying
    #geom_rect(
    #  data = subtable_info_basket,
    #  aes(xmin = xmin, xmax = xmax,
    #      ymin = ymin, ymax = ymax,
    #      fill = caregiver)
    #) +
    
    scale_fill_manual(values = caregiver_colors) +  
    
    scale_x_continuous(
      limits = c(0, 1),
      expand = c(0, 0),
      labels = scales::percent
    ) +
    scale_y_continuous(
      breaks = NULL,
      expand = c(0, 0)
    ) +
    theme_void() +
    theme(
      legend.position = "none"
    )
  # scale_y_continuous(
  #   breaks = c(y_basket, y_forage, y_forage+0.02, y_infant),
  #   #labels = c("Basket carrying", "Foraging (purple)", "Travelling (yellow)", #"Resting (green)", 
  #   #           "Infant carrying"),
  #   expand = c(0, 0)
  # ) +
  # theme_void() +
  # # labs(
  # #   fill = "Caregiver"#,
  # #   #title = paste("Caregiving, foraging and travelling –", nursing_woman)
  # # ) +
  # theme(
  #   legend.position = "bottom",
  #   axis.text.y = element_text(size = 10)
  # )
  
  
  print(p)
  
  ggsave(filename = paste(nursing_woman, "_switch.png"),
         width = 10, height = 6, dpi = 300)
  
}

```

\

## Childcare transfer statistics used in the Results

\

### With Ground

(1) the timing and frequency of childcare transfers, defined as shifts in responsibility for the infant between mothers and other group members (e.g., transitions from mother carrying to another individual carrying; C$_{mother}$ → C$_{other}$);

```{r}

#==================
# Helper: collapse consecutive duplicate caregivers
#==================
collapse_consecutive <- function(df) {
  if (nrow(df) <= 1) return(df)
  
  collapsed <- df[1, , drop = FALSE]
  
  for (row in 2:nrow(df)) {
    if (df$caregiver[row] == collapsed$caregiver[nrow(collapsed)]) {
      # Same caregiver as previous row — accumulate duration
      collapsed$duration[nrow(collapsed)] <- collapsed$duration[nrow(collapsed)] + df$duration[row]
    } else {
      # New caregiver — add a new row
      collapsed <- rbind(collapsed, df[row, , drop = FALSE])
    }
  }
  return(collapsed)
}

#==================
# of1: all transfers (including ground)
#==================
focal_follow_relay_coop$of1_childtransfer_duration <- NA
focal_follow_relay_coop$of1_childtransfer_number   <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  subset_woman <- subset(infant_caregiver, infant_caregiver$trip_id == woman)
  
  # Collapse consecutive duplicate caregivers first
  subset_woman <- collapse_consecutive(subset_woman)
  
  focal_follow_relay_coop$of1_childtransfer_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman$duration)
  
  focal_follow_relay_coop$of1_childtransfer_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman) - 1 #we do not want the number of rows but transfers!
}

summary(focal_follow_relay_coop$of1_childtransfer_number)
sd(focal_follow_relay_coop$of1_childtransfer_number)

```

\

### No Ground

```{r}

#==================
# of2: transfers excluding ground
#==================
focal_follow_relay_coop$of2_childtransfer_no_ground_duration <- NA
focal_follow_relay_coop$of2_childtransfer_no_ground_number   <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  subset_woman <- subset(infant_caregiver, infant_caregiver$trip_id == woman)
  
  # Collapse consecutive duplicate caregivers first
  subset_woman <- collapse_consecutive(subset_woman)
  
  # Then exclude ground bouts
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  
  # Collapse again in case removing ground created new consecutive duplicates
  # e.g. mother -> Ground -> mother becomes mother -> mother after ground removal
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  focal_follow_relay_coop$of2_childtransfer_no_ground_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman_no_ground$duration)
  
  focal_follow_relay_coop$of2_childtransfer_no_ground_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman_no_ground) - 1 #we do not want the number of rows but transfers!
}

summary(focal_follow_relay_coop$of2_childtransfer_no_ground_number)
sd(focal_follow_relay_coop$of2_childtransfer_no_ground_number)

```

\

### Mother and Girl

```{r}

focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl <- NA
focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother <- NA
focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  
  subset_woman <- subset(infant_caregiver, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_girl <- 0
  girl_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "girl") {
        mother_to_girl <- mother_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "mother") {
        girl_to_mother <- girl_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl
  focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_mother
  focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl + girl_to_mother
}

#summary(focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl[trip_with_girls])
#summary(focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother[trip_with_girls])

summary(focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[trip_with_girls])

```

\

## Basket-transfer statistics used in the Results

\

### Basket transfers

```{r}

### Basket transfers

#==================
# ob1: all basket transfers (including ground)
#==================
focal_follow_relay_coop$ob1_baskettransfer_duration <- NA
focal_follow_relay_coop$ob1_baskettransfer_number   <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, basket_carrier$trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  focal_follow_relay_coop$ob1_baskettransfer_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman$duration)
  
  focal_follow_relay_coop$ob1_baskettransfer_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman) - 1
}

summary(focal_follow_relay_coop$ob1_baskettransfer_duration)
summary(focal_follow_relay_coop$ob1_baskettransfer_number)
sd(focal_follow_relay_coop$ob1_baskettransfer_number, na.rm = TRUE)

#==================
# ob2: basket transfers excluding ground
#==================
focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration <- NA
focal_follow_relay_coop$ob2_baskettransfer_no_ground_number   <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, basket_carrier$trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman_no_ground$duration)
  
  focal_follow_relay_coop$ob2_baskettransfer_no_ground_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman_no_ground) - 1
}

summary(focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration)
summary(focal_follow_relay_coop$ob2_baskettransfer_no_ground_number)
sd(focal_follow_relay_coop$ob2_baskettransfer_no_ground_number, na.rm = TRUE)

#==================
# ob3/ob4: B_mother <-> B_ground
#==================
focal_follow_relay_coop$ob3_transition_Bmother_to_Bground <- NA
focal_follow_relay_coop$ob4_transition_Bground_to_Bmother <- NA
focal_follow_relay_coop$ob5_transition_Bmother_Bground_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_ground <- 0
  ground_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "Ground") {
        mother_to_ground <- mother_to_ground + 1
      }
      if (subset_woman$caregiver[i-1] == "Ground" &&
          subset_woman$caregiver[i]   == "mother") {
        ground_to_mother <- ground_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob3_transition_Bmother_to_Bground[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_ground
  focal_follow_relay_coop$ob4_transition_Bground_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_mother
  focal_follow_relay_coop$ob5_transition_Bmother_Bground_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_ground + ground_to_mother
}

summary(focal_follow_relay_coop$ob5_transition_Bmother_Bground_total)
sd(focal_follow_relay_coop$ob5_transition_Bmother_Bground_total, na.rm = TRUE)

#==================
# ob6/ob7: B_mother <-> B_other (excluding ground)
#==================
focal_follow_relay_coop$ob6_transition_Bmother_to_Bother <- NA
focal_follow_relay_coop$ob7_transition_Bother_to_Bmother <- NA
focal_follow_relay_coop$ob8_transition_Bmother_Bother_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  # Remove ground before checking mother <-> other
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  mother_to_other <- 0
  other_to_mother <- 0
  
  if (nrow(subset_woman_no_ground) > 1) {
    for (i in 2:nrow(subset_woman_no_ground)) {
      if (subset_woman_no_ground$caregiver[i-1] == "mother" &&
          subset_woman_no_ground$caregiver[i]   != "mother") {
        mother_to_other <- mother_to_other + 1
      }
      if (subset_woman_no_ground$caregiver[i-1] != "mother" &&
          subset_woman_no_ground$caregiver[i]   == "mother") {
        other_to_mother <- other_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob6_transition_Bmother_to_Bother[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_other
  focal_follow_relay_coop$ob7_transition_Bother_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    other_to_mother
  focal_follow_relay_coop$ob8_transition_Bmother_Bother_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_other + other_to_mother
}

summary(focal_follow_relay_coop$ob8_transition_Bmother_Bother_total)
sd(focal_follow_relay_coop$ob8_transition_Bmother_Bother_total, na.rm = TRUE)

#==================
# ob9/ob10: B_mother <-> B_girl
#==================
focal_follow_relay_coop$ob9_transition_Bmother_to_Bgirl   <- NA
focal_follow_relay_coop$ob10_transition_Bgirl_to_Bmother  <- NA
focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_girl <- 0
  girl_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "girl") {
        mother_to_girl <- mother_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "mother") {
        girl_to_mother <- girl_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob9_transition_Bmother_to_Bgirl[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl
  focal_follow_relay_coop$ob10_transition_Bgirl_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_mother
  focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl + girl_to_mother
}

summary(focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[trip_with_girls], na.rm = TRUE)

```

\

### Basket and girls

```{r}

#==================
# ob12/ob13: B_ground <-> B_girl
#==================
focal_follow_relay_coop$ob12_transition_Bground_to_Bgirl  <- NA
focal_follow_relay_coop$ob13_transition_Bgirl_to_Bground  <- NA
focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  ground_to_girl <- 0
  girl_to_ground <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "Ground" &&
          subset_woman$caregiver[i]   == "girl") {
        ground_to_girl <- ground_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "Ground") {
        girl_to_ground <- girl_to_ground + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob12_transition_Bground_to_Bgirl[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_girl
  focal_follow_relay_coop$ob13_transition_Bgirl_to_Bground[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_ground
  focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_girl + girl_to_ground
}

summary(focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[trip_with_girls], na.rm = TRUE)

```

\

# Figure 6: Relay cooperation percentages

This analysis is restricted to the 13 trips in which girls were present. Percentages
are calculated separately for each trip relative to the focal mother's total time
spent in the corresponding activity state (foraging, resting, or travelling).

For maternal foraging and resting, basket position is **not** included in the
childcare calculations. During travelling, the infant-care and basket-carrying
states are combined because both are part of the depicted load-sharing pattern.

**Important reproducibility note.** The four input files loaded in this script do
not contain a time-resolved measure of girls' own foraging activity. Therefore,
the manuscript statistic describing girls foraging during 57.5% of maternal
resting time cannot be regenerated from these four files alone. The source file
and code used for that statistic should be added here before archival submission.

```{r figure6-relay-percentages}

# ============================================================
# FIGURE 6: RELAY COOPERATION PERCENTAGES
# Only the 13 trips in which girls were present
# ============================================================

# ------------------------------------------------------------
# 1. Helper functions
# ------------------------------------------------------------

overlap_2way <- function(df1, df2) {

  total <- 0

  if (nrow(df1) > 0 && nrow(df2) > 0) {
    for (i in seq_len(nrow(df1))) {
      for (j in seq_len(nrow(df2))) {

        start <- max(df1$xmin[i], df2$xmin[j])
        end   <- min(df1$xmax[i], df2$xmax[j])

        total <- total + max(0, end - start)
      }
    }
  }

  total
}


overlap_3way <- function(df1, df2, df3) {

  total <- 0

  if (nrow(df1) > 0 &&
      nrow(df2) > 0 &&
      nrow(df3) > 0) {

    for (i in seq_len(nrow(df1))) {
      for (j in seq_len(nrow(df2))) {
        for (k in seq_len(nrow(df3))) {

          start <- max(
            df1$xmin[i],
            df2$xmin[j],
            df3$xmin[k]
          )

          end <- min(
            df1$xmax[i],
            df2$xmax[j],
            df3$xmax[k]
          )

          total <- total + max(0, end - start)
        }
      }
    }
  }

  total
}


# Convert HH:MM:SS to seconds
time_to_seconds <- function(x) {

  x <- strsplit(as.character(x), ":")

  sapply(x, function(z) {
    as.numeric(z[1]) * 3600 +
      as.numeric(z[2]) * 60 +
      as.numeric(z[3])
  })
}


# ------------------------------------------------------------
# 2. Identify the 13 trips with girls present
# ------------------------------------------------------------

trips_with_girls <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_girls > 0
]

print(trips_with_girls)
length(trips_with_girls)   # expected: 13


# ------------------------------------------------------------
# 3. Storage
# ------------------------------------------------------------

relay_results <- data.frame()


# ============================================================
# 4. Main loop
# ============================================================

for (woman in trips_with_girls) {

  # ==========================================================
  # A. Build maternal activity timeline directly
  #    from Waypoints_relay_coop
  # ==========================================================

  w <- Waypoints_relay_coop %>%
    filter(ID_focal == woman)

  w$time_sec <- time_to_seconds(w$Time)

  # Duration from each waypoint until the next waypoint
  w$duration <- c(
    diff(w$time_sec),
    NA
  )

  # Final row has no following interval
  w <- w[!is.na(w$duration), ]


  # ----------------------------------------------------------
  # Reconstruct travelling state
  # ----------------------------------------------------------

  walking_state <- FALSE
  w$walking_state <- FALSE

  for (i in seq_len(nrow(w))) {

    if (!is.na(w$walking[i])) {

      if (w$walking[i] == "start") {
        walking_state <- TRUE
      }

      if (w$walking[i] == "stop") {
        walking_state <- FALSE
      }
    }

    w$walking_state[i] <- walking_state
  }


  # ----------------------------------------------------------
  # Reconstruct maternal foraging state
  # ----------------------------------------------------------

  foraging_state <- FALSE
  w$foraging_state <- FALSE

  for (i in seq_len(nrow(w))) {

    if (!is.na(w$foraging[i])) {

      if (w$foraging[i] == "start") {
        foraging_state <- TRUE
      }

      if (w$foraging[i] == "stop") {
        foraging_state <- FALSE
      }
    }

    w$foraging_state[i] <- foraging_state
  }


  # resting_behaviour is already coded TRUE/FALSE
  w$resting_state <- as.logical(w$resting_behaviour)


  # ----------------------------------------------------------
  # Convert maternal intervals to fractions of the whole trip
  # ----------------------------------------------------------

  trip_duration <- sum(w$duration)

  w$frac <- w$duration / trip_duration
  w$xmax <- cumsum(w$frac)
  w$xmin <- c(0, head(w$xmax, -1))


  forage_true <- w[w$foraging_state == TRUE, ]
  rest_true   <- w[w$resting_state == TRUE, ]
  walk_true   <- w[w$walking_state == TRUE, ]


  forage_total <- sum(forage_true$frac)
  rest_total   <- sum(rest_true$frac)
  walk_total   <- sum(walk_true$frac)


  # ==========================================================
  # B. Infant-care timeline
  # ==========================================================

  infant <- infant_caregiver %>%
    filter(trip_id == woman)

  infant$frac <- infant$duration / sum(infant$duration)
  infant$xmax <- cumsum(infant$frac)
  infant$xmin <- c(0, head(infant$xmax, -1))

  mother_care <- infant %>%
    filter(caregiver == "mother")

  girl_care <- infant %>%
    filter(caregiver == "girl")


  # ==========================================================
  # C. Basket timeline
  # ==========================================================

  basket <- basket_carrier %>%
    filter(trip_id == woman)

  basket$frac <- basket$duration / sum(basket$duration)
  basket$xmax <- cumsum(basket$frac)
  basket$xmin <- c(0, head(basket$xmax, -1))

  mother_basket <- basket %>%
    filter(caregiver == "mother")

  girl_basket <- basket %>%
    filter(caregiver == "girl")


  # ==========================================================
  # 5. When mothers forage
  # Basket position is NOT included
  # ==========================================================

  overlap_F_mothercare <- overlap_2way(
    forage_true,
    mother_care
  )

  overlap_F_girlcare <- overlap_2way(
    forage_true,
    girl_care
  )


  pct_F_mothercare <- ifelse(
    forage_total > 0,
    overlap_F_mothercare / forage_total * 100,
    NA
  )

  pct_F_girlcare <- ifelse(
    forage_total > 0,
    overlap_F_girlcare / forage_total * 100,
    NA
  )


  # ==========================================================
  # 6. When mothers rest
  # ==========================================================

  overlap_R_mothercare <- overlap_2way(
    rest_true,
    mother_care
  )

  overlap_R_girlcare <- overlap_2way(
    rest_true,
    girl_care
  )


  pct_R_mothercare <- ifelse(
    rest_total > 0,
    overlap_R_mothercare / rest_total * 100,
    NA
  )

  pct_R_girlcare <- ifelse(
    rest_total > 0,
    overlap_R_girlcare / rest_total * 100,
    NA
  )


  # ==========================================================
  # 7. When mothers travel
  # ==========================================================

  # Mother carries basket + girl carries infant
  overlap_T_Mbasket_Ginfant <- overlap_3way(
    walk_true,
    mother_basket,
    girl_care
  )

  # Mother carries infant + girl carries basket
  overlap_T_Minfant_Gbasket <- overlap_3way(
    walk_true,
    mother_care,
    girl_basket
  )

  # Mother carries both infant and basket
  overlap_T_Minfant_Mbasket <- overlap_3way(
    walk_true,
    mother_care,
    mother_basket
  )


  pct_T_Mbasket_Ginfant <- ifelse(
    walk_total > 0,
    overlap_T_Mbasket_Ginfant / walk_total * 100,
    NA
  )

  pct_T_Minfant_Gbasket <- ifelse(
    walk_total > 0,
    overlap_T_Minfant_Gbasket / walk_total * 100,
    NA
  )

  pct_T_Minfant_Mbasket <- ifelse(
    walk_total > 0,
    overlap_T_Minfant_Mbasket / walk_total * 100,
    NA
  )


  # ==========================================================
  # 8. Save trip-level results
  # ==========================================================

  relay_results <- rbind(
    relay_results,
    data.frame(

      trip_id = woman,

      forage_mother_care =
        pct_F_mothercare,

      forage_girl_care =
        pct_F_girlcare,

      rest_mother_care =
        pct_R_mothercare,

      rest_girl_care =
        pct_R_girlcare,

      travel_mother_basket_girl_infant =
        pct_T_Mbasket_Ginfant,

      travel_mother_infant_girl_basket =
        pct_T_Minfant_Gbasket,

      travel_mother_infant_and_basket =
        pct_T_Minfant_Mbasket
    )
  )
}


# ============================================================
# 9. Inspect all 13 trip-level values
# ============================================================

print(relay_results)


# ============================================================
# 10. Summary function
# ============================================================

summarise_pct <- function(x) {

  data.frame(
    N = sum(!is.na(x)),
    trips_observed = sum(x > 0, na.rm = TRUE),
    mean = mean(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE)
  )
}


# ============================================================
# 11. Final Figure 6 summary
# ============================================================

figure6_summary <- rbind(

  "Mother forages + mother provides infant care" =
    summarise_pct(
      relay_results$forage_mother_care
    ),

  "Mother forages + girl provides infant care" =
    summarise_pct(
      relay_results$forage_girl_care
    ),

  "Mother rests + mother provides infant care + girl forages" =
    summarise_pct(
      relay_results$rest_mother_care
    ),

  "Mother rests + girl provides infant care" =
    summarise_pct(
      relay_results$rest_girl_care
    ),

  "Mother travels + mother carries basket + girl carries infant" =
    summarise_pct(
      relay_results$travel_mother_basket_girl_infant
    ),

  "Mother travels + mother carries infant + girl carries basket" =
    summarise_pct(
      relay_results$travel_mother_infant_girl_basket
    ),

  "Mother travels + mother carries infant and basket" =
    summarise_pct(
      relay_results$travel_mother_infant_and_basket
    )
)

round(figure6_summary, 1)
```

# Supplementary Table 3

```{r}

#==================
# Caregiver x Activity summary table
#==================

# Get all caregiver categories actually present in the data
caregiver_levels <- sort(unique(as.character(infant_caregiver$caregiver)))

# Long-format results: one row per trip x caregiver x activity
caregiver_activity_long <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {

  #------------------
  # Infant carrying
  #------------------
  infant <- subset(infant_caregiver, trip_id == nursing_woman)
  infant$frac <- infant$duration / sum(infant$duration)
  infant$xmax <- cumsum(infant$frac)
  infant$xmin <- c(0, head(infant$xmax, -1))

  #------------------
  # Foraging / Resting / Walking (same as before)
  #------------------
  forage <- subset(foraging_bout, trip_id == nursing_woman)
  forage$frac <- forage$duration / sum(forage$duration)
  forage$xmax <- cumsum(forage$frac)
  forage$xmin <- c(0, head(forage$xmax, -1))

  rest <- subset(rest_bout, trip_id == nursing_woman)
  rest$frac <- rest$duration / sum(rest$duration)
  rest$xmax <- cumsum(rest$frac)
  rest$xmin <- c(0, head(rest$xmax, -1))

  walk <- subset(walking_bout, trip_id == nursing_woman)
  walk$frac <- walk$duration / sum(walk$duration)
  walk$xmax <- cumsum(walk$frac)
  walk$xmin <- c(0, head(walk$xmax, -1))

  forage_true <- subset(forage, forages == 1)
  rest_true   <- subset(rest, rest == 1)
  walk_true   <- subset(walk, walks == 1)

  forage_total_frac <- sum(forage_true$frac)
  rest_total_frac   <- sum(rest_true$frac)
  walk_total_frac   <- sum(walk_true$frac)

  activities <- list(
    foraging   = list(bouts = forage_true, total = forage_total_frac),
    resting    = list(bouts = rest_true,   total = rest_total_frac),
    travelling = list(bouts = walk_true,   total = walk_total_frac)
  )

  #------------------
  # Loop over caregiver categories x activities
  #------------------
  for (cg in caregiver_levels) {
    cg_bouts <- subset(infant, caregiver == cg)

    for (act_name in names(activities)) {
      act_bouts <- activities[[act_name]]$bouts
      act_total <- activities[[act_name]]$total

      ov <- overlap_2way(act_bouts, cg_bouts)
      norm_val <- ifelse(act_total > 0, ov / act_total, NA)

      caregiver_activity_long <- rbind(
        caregiver_activity_long,
        data.frame(
          trip_id   = nursing_woman,
          caregiver = cg,
          activity  = act_name,
          norm_frac = norm_val
        )
      )
    }
  }
}

#==================
# Average across trips -> wide table (caregiver rows x activity columns)
#==================
caregiver_activity_summary <- aggregate(
  norm_frac ~ caregiver + activity,
  data = caregiver_activity_long,
  FUN = mean, na.rm = TRUE
)

caregiver_activity_table <- reshape(
  caregiver_activity_summary,
  idvar = "caregiver",
  timevar = "activity",
  direction = "wide"
)

# clean up column names: norm_frac.foraging -> foraging, etc.
names(caregiver_activity_table) <- gsub("norm_frac\\.", "", names(caregiver_activity_table))

# reorder columns nicely
caregiver_activity_table <- caregiver_activity_table[, c("caregiver", "foraging", "resting", "travelling")]

# convert to percentages, rounded
caregiver_activity_table$foraging   <- round(caregiver_activity_table$foraging   * 100, 1)
caregiver_activity_table$resting    <- round(caregiver_activity_table$resting    * 100, 1)
caregiver_activity_table$travelling <- round(caregiver_activity_table$travelling * 100, 1)

caregiver_activity_table

```
