---
title: "Relay Cooperation"
author: "A. Visine"
date: "2025-12-22"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_section: yes
    code_folding: show
    theme:
      bg: "white"
      fg: "#8c76c1"
      primary: "#fcc175"
      secondary: "#349a80"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

First, make sure you are saving the script and input file in the same folder, and [setting the directory to *source file*](https://stackoverflow.com/questions/13672720/r-command-for-setting-working-directory-to-source-file-location-in-rstudio).

```{r setup, include=FALSE}

rm(list = ls())  #clean environment

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/GitHub/Relay_Cooperation")

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) #set directory where the working file is at the moment
#This line can sometimes bug, closing and reopening the file solves it
#if you are not using Rstudio but Sourceing it, use the next command:
# setwd(getSrcDirectory(function(){})[1])

```

This script will, from the observational data, produce descriptive statistics used in the paper.

This file was produced using a [Rmarkdown Cheatsheet](https://rstudio.github.io/cheatsheets/html/rmarkdown.html).

\

# Libraries

This script will automatically check if the packages are installed, and if not, install and launch them.

```{r}

# Helper function: install if missing, then load
load_or_install <- function(pkg){
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
  library(pkg, character.only = TRUE)
}

# Packages you need
packages <- c("readr",  #read video_data file
              "ggplot2")#plot data

# Install & load all
lapply(packages, load_or_install)

```

\

# Load

This script loads the data from focal follows, with the files a) *Waypoints_relay_coop*, a csv file containing for each row a change of activity in mother's activities, childcare responsibilities or basket carrying.  

This csv contains the columns:  

1. Focal follow / focal woman identifier
    - ID_focal:           focal woman and trip ID; num. from 1 to 18

2. Focal woman activity
    - Time:               hour::minute::second of the day when the focal woman **starts** a different behaviour/ activity
    - resting_behaviour:  is the focal woman resting? log. TRUE/FALSE
    - walking:            is the focal woman travelling? log. TRUE/FALSE
    - foraging:           is the focal woman foraging? log. TRUE/FALSE
    - labour:             is the focal woman working? foraging is included in this measure as well log. TRUE/FALSE
    - breastfeed:         did the focal woman start breastfeeding? log. TRUE/FALSE

\

We also load b) *focal_follow_relay_coop*, which describes and summarises each focal follow independently (*i.e.*, one row per focal follow = 18 rows). When loaded, it contains the columns:

1. Focal follow / focal woman identifier
    - ID_focal:           focal woman and trip ID; num. from 1 to 18
    - trip_duration:      total duration of the foraging trip in seconds

2. Group composition
    - n_girls:            total number of girls present during the foraging trip
    - n_women:            total number of women present during the foraging trip
    - n_boys:             total number of boys present during the foraging trip
    - n_men:              total number of men present during the foraging trip
    - husband_presence:   presence of the focal woman's husband (and infant father); log. TRUE/FALSE

\

Then we load c) *infant_caregiver*, a table describing the infant caregiver at any time of the trip. It has 3 columns:
  - trip_id:      focal woman and trip ID; num. from 1 to 18
- caregiver:    identity of the infant caregiver at a given time (mother or girl or woman or boy or father or ground or unknown)
-duration:      duration of the caregiving configuration in seconds (cumulative duration for each ID equals to the total trip duration)

\

Finally we load c) *basket_carrier*, a table describing the focal woman's basket carrier at any time of the trip. It has the same 3 columns:
    - trip_id:      focal woman and trip ID; num. from 1 to 18
    - caregiver:    identity of the basket carrier at a given time (mother or girl or woman or boy or father or ground or unknown)
    -duration:      duration of the carrying configuration in seconds (cumulative duration for each ID equals to the total trip duration)

\

```{r}

#load(".RData") #load the R environment

Waypoints_relay_coop <- read_delim("Waypoints_relay_coop.csv", col_names = T, delim = ",")

focal_follow_relay_coop <- read_delim("focal_follow_relay_coop.csv", col_names = T, delim = ",")

infant_caregiver <- read_delim("infant_caregiver.csv", col_names = T, delim = ",")

basket_carrier <- read_delim("basket_carrier.csv", col_names = T, delim = ",")

```

\


# Transformation

## Trips with girls

```{r}

#List of trips where girls are present (13 trips)

trip_with_girls <- which(focal_follow_relay_coop$n_girls > 0)

```

\

## Bouts tables

These tables we be used to produce figure 5.

\

### Resting bouts

```{r}

rest_bout <- data.frame(
    trip_id   = c("TEST"),
    rest = c("TEST"),
    duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  for (step_rest in 2:nrow(nursing_w)-1) {
    
    rest_bout <- rbind(rest_bout, c(woman,
                                        nursing_w$resting_behaviour[step_rest],
                                        nursing_w$Time[step_rest+1] - nursing_w$Time[step_rest]))
  
    
  }
}

#Delete TEST row
rest_bout <- rest_bout[-1,]

#Make sure it is numerical now
rest_bout$duration <- as.numeric(rest_bout$duration)

#Make sure it is logical now
#rest_bout$rest <- as.logical.factor(rest_bout$duration)


```

\

### Foraging bouts

- Foraging

```{r}

foraging_bout <- data.frame(
    trip_id   = c("TEST"),
    forages = c("TEST"),
    duration = c("TEST"))

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  #Keep one woman after the other
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == nursing_woman)
  
  #Keep the first and last row, to know when the trip is ending, for duration calculation
  first_row <- nursing_w[1,]
  last_row <- nursing_w[nrow(nursing_w),]
  
  #Keep only changes in forages
  nursing_w <- subset(nursing_w, is.na(nursing_w$foraging) == FALSE)
  
  #Add back the last row
  nursing_w <- rbind(first_row, nursing_w, last_row)
  
  #First bout where no foraging
  foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[2] - nursing_w$Time[1]))
  
  #For each change in foraging
  for (change in 4:nrow(nursing_w)-2) {
    #If the mother forages, add TRUE in the bout file
    if(nursing_w$foraging[change] == "start"){
      foraging_bout <- rbind(foraging_bout, c(nursing_woman, TRUE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    } else { #if there is no forages, add their name, we'll replace by boy/girl/etc later
foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[change+1] - nursing_w$Time[change]))
}
}

#Last bout where no foraging
foraging_bout <- rbind(foraging_bout, c(nursing_woman, FALSE, nursing_w$Time[nrow(nursing_w)] - nursing_w$Time[nrow(nursing_w)-1]))

}  


#Delete the TEST row
foraging_bout <- foraging_bout[-1,]

#Make sure it is numerical now
foraging_bout$duration <- as.numeric(foraging_bout$duration)


```

\

### Walking bouts

- Waking

```{r}

walking_bout <- data.frame(
  trip_id   = c("TEST"),
  walks = c("TEST"),
  duration = c("TEST"))

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  #Keep one woman after the other
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == nursing_woman)
  
  
  #Keep only changes in walking
  nursing_w <- subset(nursing_w, is.na(nursing_w$walking) == FALSE)
  
  
  #For each change in walking
  for (change in 2:nrow(nursing_w)-1) {
    #If the mother walks, add TRUE in the bout file
    if(nursing_w$walking[change] == "start"){
      walking_bout <- rbind(walking_bout, c(nursing_woman, TRUE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    } else { #if there is no walks, add their name, we'll replace by boy/girl/etc later
      walking_bout <- rbind(walking_bout, c(nursing_woman, FALSE, nursing_w$Time[change+1] - nursing_w$Time[change]))
    }
  }
  
}  

#Delete the TEST row
walking_bout <- walking_bout[-1,]

#Make sure it is numerical now
walking_bout$duration <- as.numeric(walking_bout$duration)


```

\

### Labour bouts

```{r}

labour_bout <- data.frame(
  trip_id   = c("TEST"),
  labour = c("TEST"),
  duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  for (step_labour in 2:nrow(nursing_w)-1) {
    
    labour_bout <- rbind(labour_bout, c(woman,
                                        nursing_w$labour[step_labour],
                                        nursing_w$Time[step_labour+1] - nursing_w$Time[step_labour]))
  }
}


#Delete the TEST row
labour_bout <- labour_bout[-1,]

#Make sure it is numerical now
labour_bout$duration <- as.numeric(labour_bout$duration)


```

\

### Breastfeeding

```{r}

#Although this is coded as "bouts" as it was before, the duration obtained corresponds to the START of breastfeeding behaviour from the beginning of the trip

breastfeeding_bouts <- data.frame(
  trip_id   = c("TEST"),
  breastfeed = c("TEST"),
  duration = c("TEST"))

for (woman in focal_follow_relay_coop$ID_focal) {
  nursing_w <- subset(Waypoints_relay_coop, Waypoints_relay_coop$ID_focal == woman)
  
  time_initial <- nursing_w$Time[1]
  
  for (step_breast in 2:nrow(nursing_w)-1) {
    
    if(nursing_w$breastfeed[step_breast] == TRUE){
      
      breastfeeding_bouts <- rbind(breastfeeding_bouts, c(woman,
                                                          TRUE,
                                                          nursing_w$Time[step_breast] - time_initial))
    }
  }
}

breastfeeding_bouts <- breastfeeding_bouts[-1,]

#Make sure it is numerical now
breastfeeding_bouts$duration <- as.numeric(breastfeeding_bouts$duration)



```


### Verification

```{r}

#Test, all ok

stopifnot(sum(walking_bout$duration, na.rm = TRUE) == sum(labour_bout$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(infant_caregiver$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(foraging_bout$duration, na.rm = TRUE),
          sum(walking_bout$duration, na.rm = TRUE) == sum(rest_bout$duration, na.rm = TRUE))

```


# Stats & Plots

## Figure 5

The plots:
  
  ```{r}

caregiver_colors <- c(
  # "boy"    = "#d29485",
  "girl"   = "#85d285",
  #"woman"  = "#85d2b3",
  "other"  = "grey",
  "mother" = "#85abd2",
  "Ground" = "#d2ab85"
)

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  subtable_info <- subset(infant_caregiver, infant_caregiver$trip_id == nursing_woman)
  subtable_info_basket <- subset(basket_carrier, basket_carrier$trip_id == nursing_woman)
  subtable_info_forage <- subset(foraging_bout, foraging_bout$trip_id == nursing_woman)
  subtable_walking <- subset(walking_bout, trip_id == nursing_woman)
  subtable_labour <-  subset(labour_bout, trip_id == nursing_woman)
  subtable_rest <-  subset(rest_bout, trip_id == nursing_woman)
  subtable_breastfeed <- subset(breastfeeding_bouts, trip_id == nursing_woman & breastfeed == 1)
  
  
  # Compute cumulative positions
  subtable_info$frac <- subtable_info$duration / sum(subtable_info$duration)
  subtable_info$xmax <- cumsum(subtable_info$frac)
  subtable_info$xmin <- c(0, head(subtable_info$xmax, -1))
  
  subtable_info_basket$frac <- subtable_info_basket$duration / sum(subtable_info_basket$duration)
  subtable_info_basket$xmax <- cumsum(subtable_info_basket$frac)
  subtable_info_basket$xmin <- c(0, head(subtable_info_basket$xmax, -1))
  
  subtable_info_forage$frac <- subtable_info_forage$duration / sum(subtable_info_forage$duration)
  subtable_info_forage$xmax <- cumsum(subtable_info_forage$frac)
  subtable_info_forage$xmin <- c(0, head(subtable_info_forage$xmax, -1))
  
  subtable_walking$frac <- subtable_walking$duration / sum(subtable_walking$duration)
  subtable_walking$xmax <- cumsum(subtable_walking$frac)
  subtable_walking$xmin <- c(0, head(subtable_walking$xmax, -1))
  
  subtable_labour$frac <- subtable_labour$duration / sum(subtable_labour$duration)
  subtable_labour$xmax <- cumsum(subtable_labour$frac)
  subtable_labour$xmin <- c(0, head(subtable_labour$xmax, -1))
  
  subtable_rest$frac <- subtable_rest$duration / sum(subtable_rest$duration)
  subtable_rest$xmax <- cumsum(subtable_rest$frac)
  subtable_rest$xmin <- c(0, head(subtable_rest$xmax, -1))
  
  subtable_breastfeed$x <- subtable_breastfeed$duration / sum(subtable_info$duration)
  
  # Keep only walking and foraging TRUE
  subtable_walking <- subset(subtable_walking, walks == 1)
  subtable_info_forage <- subset(subtable_info_forage, subtable_info_forage$forages == 1)
  subtable_labour <- subset(subtable_labour, labour == 1)
  subtable_rest <- subset(subtable_rest, rest == 1)
  
  #Keep only mother and girls
  #subtable_info <- subset(subtable_info, subtable_info$caregiver %in% c("mother", "girl", "Ground"))
  #subtable_info_basket <- subset(subtable_info_basket, subtable_info_basket$caregiver %in% c("mother", "girl", "Ground"))
  
  #Assign bar position
  bar_height <- 0.05
  
  y_infant   <-  0.20
  y_forage   <-  0.10
  y_basket   <-  0.00
  
  
  subtable_info$ymin <- y_infant - bar_height
  subtable_info$ymax <- y_infant + bar_height
  
  subtable_info_forage$ymin <- y_forage - bar_height
  subtable_info_forage$ymax <- y_forage + bar_height
  
  subtable_walking$ymin <- y_forage - bar_height
  subtable_walking$ymax <- y_forage + bar_height
  
  subtable_labour$ymin <- numeric(nrow(subtable_labour))
  subtable_labour$ymax <- numeric(nrow(subtable_labour))
  
  if (nrow(subtable_labour) > 0) {
    subtable_labour$ymin <- y_forage - bar_height
    subtable_labour$ymax <- y_forage + bar_height
  }
  
  subtable_rest$ymin <- numeric(nrow(subtable_rest))
  subtable_rest$ymax <- numeric(nrow(subtable_rest))
  
  if (nrow(subtable_rest) > 0) {
    subtable_rest$ymin <- y_forage - bar_height
    subtable_rest$ymax <- y_forage + bar_height
  }
  
  subtable_info_basket$ymin <- y_basket - bar_height
  subtable_info_basket$ymax <- y_basket + bar_height
  
  
  # Plot
  p <- ggplot() +
    
    # Infant holding
    geom_rect(
      data = subtable_info,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax,
          fill = caregiver)
    ) +
    
    geom_segment(
      data = subtable_breastfeed,
      aes(
        x = x, xend = x,
        y = y_infant - bar_height,
        yend = y_infant + bar_height
      ),
      colour = "darkblue",
      linetype = "dotted",
      linewidth = 0.8
    ) +
    
    # Walking (overlay on same bar)
    geom_rect(
      data = subtable_walking,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#ffd690",
      alpha = 1
    ) +
    
    # Rest
    geom_rect(
      data = subtable_rest,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#b8dcb8",
      alpha = 1
    ) +
    
    # Labour (background bar)
    geom_rect(
      data = subtable_labour,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#c3b0d5",
      alpha = 1
    ) +
    
    # Foraging
    geom_rect(
      data = subtable_info_forage,
      aes(xmin = xmin, xmax = xmax,
          ymin = ymin, ymax = ymax),
      fill = "#756a80",
      alpha = 1
    ) +
    
    # Basket carrying
    #geom_rect(
    #  data = subtable_info_basket,
    #  aes(xmin = xmin, xmax = xmax,
    #      ymin = ymin, ymax = ymax,
    #      fill = caregiver)
    #) +
    
    scale_fill_manual(values = caregiver_colors) +  
    
    scale_x_continuous(
      limits = c(0, 1),
      expand = c(0, 0),
      labels = scales::percent
    ) +
    scale_y_continuous(
      breaks = NULL,
      expand = c(0, 0)
    ) +
    theme_void() +
    theme(
      legend.position = "none"
    )
  # scale_y_continuous(
  #   breaks = c(y_basket, y_forage, y_forage+0.02, y_infant),
  #   #labels = c("Basket carrying", "Foraging (purple)", "Travelling (yellow)", #"Resting (green)", 
  #   #           "Infant carrying"),
  #   expand = c(0, 0)
  # ) +
  # theme_void() +
  # # labs(
  # #   fill = "Caregiver"#,
  # #   #title = paste("Caregiving, foraging and travelling –", nursing_woman)
  # # ) +
  # theme(
  #   legend.position = "bottom",
  #   axis.text.y = element_text(size = 10)
  # )
  
  
  print(p)
  
  ggsave(filename = paste(nursing_woman, "_switch.png"),
         width = 10, height = 6, dpi = 300)
  
}

```

\

## *Women foraging and girls providing childcare*, Childcare transferts

\

### With Ground

(1) the timing and frequency of childcare transfers, defined as shifts in responsibility for the infant between mothers and other group members (e.g., transitions from mother carrying to another individual carrying; C$_{mother}$ → C$_{other}$);

```{r}

#==================
# Helper: collapse consecutive duplicate caregivers
#==================
collapse_consecutive <- function(df) {
  if (nrow(df) <= 1) return(df)
  
  collapsed <- df[1, , drop = FALSE]
  
  for (row in 2:nrow(df)) {
    if (df$caregiver[row] == collapsed$caregiver[nrow(collapsed)]) {
      # Same caregiver as previous row — accumulate duration
      collapsed$duration[nrow(collapsed)] <- collapsed$duration[nrow(collapsed)] + df$duration[row]
    } else {
      # New caregiver — add a new row
      collapsed <- rbind(collapsed, df[row, , drop = FALSE])
    }
  }
  return(collapsed)
}

#==================
# of1: all transfers (including ground)
#==================
focal_follow_relay_coop$of1_childtransfer_duration <- NA
focal_follow_relay_coop$of1_childtransfer_number   <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  subset_woman <- subset(infant_caregiver, infant_caregiver$trip_id == woman)
  
  # Collapse consecutive duplicate caregivers first
  subset_woman <- collapse_consecutive(subset_woman)
  
  focal_follow_relay_coop$of1_childtransfer_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman$duration)
  
  focal_follow_relay_coop$of1_childtransfer_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman) - 1 #we do not want the number of rows but transfers!
}

#focal_follow_relay_coop$of1_childtransfer_duration
summary(focal_follow_relay_coop$of1_childtransfer_number)
sd(focal_follow_relay_coop$of1_childtransfer_number)

```

\

### No Ground

```{r}

#==================
# of2: transfers excluding ground
#==================
focal_follow_relay_coop$of2_childtransfer_no_ground_duration <- NA
focal_follow_relay_coop$of2_childtransfer_no_ground_number   <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  subset_woman <- subset(infant_caregiver, infant_caregiver$trip_id == woman)
  
  # Collapse consecutive duplicate caregivers first
  subset_woman <- collapse_consecutive(subset_woman)
  
  # Then exclude ground bouts
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  
  # Collapse again in case removing ground created new consecutive duplicates
  # e.g. mother -> Ground -> mother becomes mother -> mother after ground removal
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  focal_follow_relay_coop$of2_childtransfer_no_ground_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman_no_ground$duration)
  
  focal_follow_relay_coop$of2_childtransfer_no_ground_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman_no_ground) - 1 #we do not want the number of rows but transfers!
}

#focal_follow_relay_coop$of2_childtransfer_no_ground_duration
summary(focal_follow_relay_coop$of2_childtransfer_no_ground_number)
sd(focal_follow_relay_coop$of2_childtransfer_no_ground_number)

```

\

### Mother and Girl

```{r}

focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl <- NA
focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother <- NA
focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  
  subset_woman <- subset(infant_caregiver, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_girl <- 0
  girl_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "girl") {
        mother_to_girl <- mother_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "mother") {
        girl_to_mother <- girl_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl
  focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_mother
  focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl + girl_to_mother
}

#summary(focal_follow_relay_coop$of5_transition_Cmother_to_Cgirl[trip_with_girls])
#summary(focal_follow_relay_coop$of6_transition_Cgirl_to_Cmother[trip_with_girls])

summary(focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$of7_transition_Cmother_Cgirl_total[trip_with_girls])

```

\

### Foraging and childcare

```{r}

focal_follow_relay_coop$of3_Fmother_Cmother_duration   <- NA
focal_follow_relay_coop$of3_Fmother_Cmother_number     <- NA
focal_follow_relay_coop$of4_Fmother_Cgirl_duration     <- NA
focal_follow_relay_coop$of4_Fmother_Cgirl_number       <- NA
focal_follow_relay_coop$of5_transition_FmCm_to_FmCg   <- NA

for (woman in unique(infant_caregiver$trip_id[-1])) {
  
  #------------------
  # Infant caregiver bouts — collapsed and normalised
  #------------------
  infant <- subset(infant_caregiver, trip_id == woman)
  infant <- collapse_consecutive(infant)
  infant$frac <- infant$duration / sum(infant$duration)
  infant$xmax <- cumsum(infant$frac)
  infant$xmin <- c(0, head(infant$xmax, -1))
  
  #------------------
  # Foraging bouts — normalised
  #------------------
  forage <- subset(foraging_bout, trip_id == woman)
  forage$frac <- forage$duration / sum(forage$duration)
  forage$xmax <- cumsum(forage$frac)
  forage$xmin <- c(0, head(forage$xmax, -1))
  forage_true <- subset(forage, forages == TRUE)
  
  #------------------
  # Tag each infant caregiver bout: was mother foraging during it?
  #------------------
  infant$mother_foraging <- FALSE
  if (nrow(forage_true) > 0) {
    for (i in seq_len(nrow(infant))) {
      for (j in seq_len(nrow(forage_true))) {
        overlap <- max(0, min(infant$xmax[i], forage_true$xmax[j]) -
                         max(infant$xmin[i], forage_true$xmin[j]))
        if (overlap > 0) {
          infant$mother_foraging[i] <- TRUE
        }
      }
    }
  }
  
  #------------------
  # Combined state per bout
  #------------------
  infant$state <- paste0(
    ifelse(infant$mother_foraging, "F_mother", "notF"),
    "_",
    ifelse(infant$caregiver == "mother", "C_mother",
           ifelse(infant$caregiver == "girl",   "C_girl",
                  paste0("C_", infant$caregiver)))
  )
  
  #------------------
  # F_mother + C_mother bouts
  #------------------
  Fm_Cm <- subset(infant, state == "F_mother_C_mother")
  focal_follow_relay_coop$of3_Fmother_Cmother_duration[focal_follow_relay_coop$ID_focal == woman] <-
    sum(Fm_Cm$duration)
  focal_follow_relay_coop$of3_Fmother_Cmother_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(Fm_Cm)
  
  #------------------
  # F_mother + C_girl bouts
  #------------------
  Fm_Cg <- subset(infant, state == "F_mother_C_girl")
  focal_follow_relay_coop$of4_Fmother_Cgirl_duration[focal_follow_relay_coop$ID_focal == woman] <-
    sum(Fm_Cg$duration)
  focal_follow_relay_coop$of4_Fmother_Cgirl_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(Fm_Cg)
  
  #------------------
  # Transitions: F_mother+C_mother --> F_mother+C_girl
  #------------------
  transitions <- 0
  if (nrow(infant) > 1) {
    for (i in 2:nrow(infant)) {
      if (infant$state[i-1] == "F_mother_C_mother" &&
          infant$state[i]   == "F_mother_C_girl") {
        transitions <- transitions + 1
      }
    }
  }
  focal_follow_relay_coop$of5_transition_FmCm_to_FmCg[focal_follow_relay_coop$ID_focal == woman] <-
    transitions
}

#------------------
# Display results
#------------------
cat("=== F_mother + C_mother ===\n")
cat("Duration (mean across trips):\n")
print(summary(focal_follow_relay_coop$of3_Fmother_Cmother_duration))
cat("Number of bouts (mean across trips):\n")
print(summary(focal_follow_relay_coop$of3_Fmother_Cmother_number))

cat("\n=== F_mother + C_girl ===\n")
cat("Duration (mean across trips):\n")
print(summary(focal_follow_relay_coop$of4_Fmother_Cgirl_duration[trip_with_girls])) # of trips with girls!
cat("Number of bouts (mean across trips):\n")
print(summary(focal_follow_relay_coop$of4_Fmother_Cgirl_number[trip_with_girls])) # of trips with girls!

cat("\n=== Transitions: F_mother+C_mother --> F_mother+C_girl ===\n")
print(summary(focal_follow_relay_coop$of5_transition_FmCm_to_FmCg[trip_with_girls]))
cat("Trips with at least one such transition:", 
    sum(focal_follow_relay_coop$of5_transition_FmCm_to_FmCg > 0, na.rm = TRUE),
    "/",
    sum(!is.na(focal_follow_relay_coop$of5_transition_FmCm_to_FmCg[trip_with_girls])), "\n")

cat("\n=== Per trip overview ===\n")
print(focal_follow_relay_coop[, c("ID_focal",
                                  "of3_Fmother_Cmother_duration",
                                  "of3_Fmother_Cmother_number",
                                  "of4_Fmother_Cgirl_duration",
                                  "of4_Fmother_Cgirl_number",
                                  "of5_transition_FmCm_to_FmCg")])

```

\

## *Infant and basket carrying* transfers

\

### Basket transfers

```{r}

### Basket transfers

#==================
# ob1: all basket transfers (including ground)
#==================
focal_follow_relay_coop$ob1_baskettransfer_duration <- NA
focal_follow_relay_coop$ob1_baskettransfer_number   <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, basket_carrier$trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  focal_follow_relay_coop$ob1_baskettransfer_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman$duration)
  
  focal_follow_relay_coop$ob1_baskettransfer_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman) - 1
}

summary(focal_follow_relay_coop$ob1_baskettransfer_duration)
summary(focal_follow_relay_coop$ob1_baskettransfer_number)
sd(focal_follow_relay_coop$ob1_baskettransfer_number, na.rm = TRUE)

#==================
# ob2: basket transfers excluding ground
#==================
focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration <- NA
focal_follow_relay_coop$ob2_baskettransfer_no_ground_number   <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, basket_carrier$trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration[focal_follow_relay_coop$ID_focal == woman] <-
    mean(subset_woman_no_ground$duration)
  
  focal_follow_relay_coop$ob2_baskettransfer_no_ground_number[focal_follow_relay_coop$ID_focal == woman] <-
    nrow(subset_woman_no_ground) - 1
}

summary(focal_follow_relay_coop$ob2_baskettransfer_no_ground_duration)
summary(focal_follow_relay_coop$ob2_baskettransfer_no_ground_number)
sd(focal_follow_relay_coop$ob2_baskettransfer_no_ground_number, na.rm = TRUE)

#==================
# ob3/ob4: B_mother <-> B_ground
#==================
focal_follow_relay_coop$ob3_transition_Bmother_to_Bground <- NA
focal_follow_relay_coop$ob4_transition_Bground_to_Bmother <- NA
focal_follow_relay_coop$ob5_transition_Bmother_Bground_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_ground <- 0
  ground_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "Ground") {
        mother_to_ground <- mother_to_ground + 1
      }
      if (subset_woman$caregiver[i-1] == "Ground" &&
          subset_woman$caregiver[i]   == "mother") {
        ground_to_mother <- ground_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob3_transition_Bmother_to_Bground[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_ground
  focal_follow_relay_coop$ob4_transition_Bground_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_mother
  focal_follow_relay_coop$ob5_transition_Bmother_Bground_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_ground + ground_to_mother
}
# 
# summary(focal_follow_relay_coop$ob3_transition_Bmother_to_Bground)
# summary(focal_follow_relay_coop$ob4_transition_Bground_to_Bmother)
summary(focal_follow_relay_coop$ob5_transition_Bmother_Bground_total)
sd(focal_follow_relay_coop$ob5_transition_Bmother_Bground_total, na.rm = TRUE)

#==================
# ob6/ob7: B_mother <-> B_other (excluding ground)
#==================
focal_follow_relay_coop$ob6_transition_Bmother_to_Bother <- NA
focal_follow_relay_coop$ob7_transition_Bother_to_Bmother <- NA
focal_follow_relay_coop$ob8_transition_Bmother_Bother_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  # Remove ground before checking mother <-> other
  subset_woman_no_ground <- subset(subset_woman, caregiver != "Ground")
  subset_woman_no_ground <- collapse_consecutive(subset_woman_no_ground)
  
  mother_to_other <- 0
  other_to_mother <- 0
  
  if (nrow(subset_woman_no_ground) > 1) {
    for (i in 2:nrow(subset_woman_no_ground)) {
      if (subset_woman_no_ground$caregiver[i-1] == "mother" &&
          subset_woman_no_ground$caregiver[i]   != "mother") {
        mother_to_other <- mother_to_other + 1
      }
      if (subset_woman_no_ground$caregiver[i-1] != "mother" &&
          subset_woman_no_ground$caregiver[i]   == "mother") {
        other_to_mother <- other_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob6_transition_Bmother_to_Bother[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_other
  focal_follow_relay_coop$ob7_transition_Bother_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    other_to_mother
  focal_follow_relay_coop$ob8_transition_Bmother_Bother_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_other + other_to_mother
}

# summary(focal_follow_relay_coop$ob6_transition_Bmother_to_Bother)
# summary(focal_follow_relay_coop$ob7_transition_Bother_to_Bmother)
summary(focal_follow_relay_coop$ob8_transition_Bmother_Bother_total)
sd(focal_follow_relay_coop$ob8_transition_Bmother_Bother_total, na.rm = TRUE)

#==================
# ob9/ob10: B_mother <-> B_girl
#==================
focal_follow_relay_coop$ob9_transition_Bmother_to_Bgirl   <- NA
focal_follow_relay_coop$ob10_transition_Bgirl_to_Bmother  <- NA
focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  mother_to_girl <- 0
  girl_to_mother <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "mother" &&
          subset_woman$caregiver[i]   == "girl") {
        mother_to_girl <- mother_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "mother") {
        girl_to_mother <- girl_to_mother + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob9_transition_Bmother_to_Bgirl[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl
  focal_follow_relay_coop$ob10_transition_Bgirl_to_Bmother[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_mother
  focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    mother_to_girl + girl_to_mother
}

# summary(focal_follow_relay_coop$ob9_transition_Bmother_to_Bgirl)
# summary(focal_follow_relay_coop$ob10_transition_Bgirl_to_Bmother)
summary(focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$ob11_transition_Bmother_Bgirl_total[trip_with_girls], na.rm = TRUE)

```

\

### Basket and girls

```{r}

#==================
# ob12/ob13: B_ground <-> B_girl
#==================
focal_follow_relay_coop$ob12_transition_Bground_to_Bgirl  <- NA
focal_follow_relay_coop$ob13_transition_Bgirl_to_Bground  <- NA
focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total <- NA

for (woman in unique(basket_carrier$trip_id[-1])) {
  subset_woman <- subset(basket_carrier, trip_id == woman)
  subset_woman <- collapse_consecutive(subset_woman)
  
  ground_to_girl <- 0
  girl_to_ground <- 0
  
  if (nrow(subset_woman) > 1) {
    for (i in 2:nrow(subset_woman)) {
      if (subset_woman$caregiver[i-1] == "Ground" &&
          subset_woman$caregiver[i]   == "girl") {
        ground_to_girl <- ground_to_girl + 1
      }
      if (subset_woman$caregiver[i-1] == "girl" &&
          subset_woman$caregiver[i]   == "Ground") {
        girl_to_ground <- girl_to_ground + 1
      }
    }
  }
  
  focal_follow_relay_coop$ob12_transition_Bground_to_Bgirl[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_girl
  focal_follow_relay_coop$ob13_transition_Bgirl_to_Bground[focal_follow_relay_coop$ID_focal == woman] <-
    girl_to_ground
  focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[focal_follow_relay_coop$ID_focal == woman] <-
    ground_to_girl + girl_to_ground
}

# summary(focal_follow_relay_coop$ob12_transition_Bground_to_Bgirl)
# summary(focal_follow_relay_coop$ob13_transition_Bgirl_to_Bground)
summary(focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[trip_with_girls])
sd(focal_follow_relay_coop$ob14_transition_Bground_Bgirl_total[trip_with_girls], na.rm = TRUE)

```

\

## Figure 4

\

### Childcare duration

```{r}

childcare_duration <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  infant <- subset(infant_caregiver,
                   trip_id == nursing_woman)
  
  mother_duration <- sum(infant$duration[infant$caregiver == "mother"], na.rm = TRUE)
  ground_duration  <- sum(infant$duration[infant$caregiver == "Ground"], na.rm = TRUE)
  other_duration  <- sum(infant$duration[!infant$caregiver %in% c("mother", "Ground")], na.rm = TRUE)
  
  childcare_duration <- rbind(
    childcare_duration,
    data.frame(
      trip_id         = nursing_woman,
      mother_duration = mother_duration,
      ground_duration = ground_duration,
      other_duration  = other_duration
    )
  )
}

childcare_duration
childcare_duration[trip_with_girls,]

```

```{r}

childcare_duration_detailed <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  infant <- subset(infant_caregiver,
                   trip_id == nursing_woman)
  
  by_caregiver <- aggregate(duration ~ caregiver, data = infant, sum)
  by_caregiver$trip_id <- nursing_woman
  
  childcare_duration_detailed <- rbind(childcare_duration_detailed, by_caregiver)
}

childcare_duration_detailed

```

```{r}

caregiver_colours <- c(
  "mother"  = "#85abd2",
  "woman"   = "#85d2b9",
  "girl"    = "#85d270",
  "boy"     = "#d29485",
  "father"  = "#d294b3",
  "ground"  = "#d2ab85",
  "unknown" = "grey"
)

childcare_frac <- childcare_duration_detailed %>%
  group_by(trip_id) %>%
  mutate(frac = duration / sum(duration)) %>%
  ungroup()

# Rename ? to unknown
childcare_frac$caregiver[childcare_frac$caregiver %in% c("Ground", "?")] <- "ground"

# Set factor order (bottom to top in stacked bar)
childcare_frac$caregiver <- factor(
  childcare_frac$caregiver,
  levels = c("ground", "father", "boy", "girl", "woman", "mother")
)

# Assign trip numbers 1 to 18 in the order they appear
childcare_frac$trip_number <- as.numeric(factor(childcare_frac$trip_id,
                                                levels = unique(childcare_frac$trip_id)))
childcare_frac$trip_number <- factor(childcare_frac$trip_number)

ggplot(childcare_frac,
       aes(x = trip_number, y = frac, fill = caregiver)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = caregiver_colours) +
  labs(
    x = "Focal trip",
    y = "Proportion of trip duration",
    fill = "Category",
    title = "A. Infant care"
  ) +
  theme_minimal() +
  theme(# Plot title
    plot.title = element_text(size = 16),
    
    # Axis titles
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    
    # Axis tick labels (numbers/text)
    axis.text.x = element_text(size = 16, angle = 0, hjust = 1),
    axis.text.y = element_text(size = 16),
    
    # Legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

ggsave("duration_prop_childcare.pdf",
       width = 10, height = 6, dpi = 300)

```

\

```{r}

#Just mother percentage
summary(c(childcare_frac$frac[which(childcare_frac$caregiver == "mother")], 0))
sd(c(childcare_frac$frac[which(childcare_frac$caregiver == "mother")], 0))

```

\

### Basket duration

```{r}

basket_duration_detailed <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  basket <- subset(basket_carrier,
                   trip_id == nursing_woman)
  
  by_caregiver <- aggregate(duration ~ caregiver, data = basket, sum)
  by_caregiver$trip_id <- nursing_woman
  
  basket_duration_detailed <- rbind(basket_duration_detailed, by_caregiver)
}

basket_duration_detailed

```

```{r}

basket_frac <- basket_duration_detailed %>%
  group_by(trip_id) %>%
  mutate(frac = duration / sum(duration)) %>%
  ungroup()

# Rename to match colour vector
basket_frac$caregiver[basket_frac$caregiver %in% c("Ground", "?")] <- "ground"

# Set factor order (bottom to top in stacked bar)
basket_frac$caregiver <- factor(
  basket_frac$caregiver,
  levels = c("ground", "father", "boy", "girl", "woman", "mother")
)

# Assign trip numbers 1 to 18 in the order they appear
basket_frac$trip_number <- as.numeric(factor(basket_frac$trip_id,
                                             levels = unique(basket_frac$trip_id)))
basket_frac$trip_number <- factor(basket_frac$trip_number)

ggplot(basket_frac,
       aes(x = trip_number, y = frac, fill = caregiver)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = caregiver_colours) +
  labs(
    x = "Focal trip",
    y = "Proportion of trip duration",
    fill = "Category",
    title = "B. Basket carrying",
    size = 16
  ) +
  theme_minimal() +
  theme(# Plot title
    plot.title = element_text(size = 16),
    
    # Axis titles
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    
    # Axis tick labels (numbers/text)
    axis.text.x = element_text(size = 16, angle = 0, hjust = 1),
    axis.text.y = element_text(size = 16),
    
    # Legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16)
  )

ggsave("duration_prop_basket.pdf",
       width = 10, height = 6, dpi = 300)

```

\

```{r}

#Just mother percentage
summary(basket_frac$frac[which(basket_frac$caregiver == "mother")])
sd(basket_frac$frac[which(basket_frac$caregiver == "mother")])

#Just ground percentage
summary(basket_frac$frac[which(basket_frac$caregiver == "ground")])
sd(basket_frac$frac[which(basket_frac$caregiver == "ground")])

```

\

### On the ground

```{r}

basket_floor_pct <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  
  basket <- subset(basket_carrier,
                   trip_id == nursing_woman)
  
  total_duration <- sum(basket$duration, na.rm = TRUE)
  floor_duration <- sum(basket$duration[basket$caregiver == "Ground"], na.rm = TRUE)
  
  basket_floor_pct <- rbind(
    basket_floor_pct,
    data.frame(
      trip_id    = nursing_woman,
      floor_duration = floor_duration,
      total_duration = total_duration,
      floor_pct  = floor_duration / total_duration * 100
    )
  )
}

basket_floor_pct
summary(basket_floor_pct$floor_pct)
sd(basket_floor_pct$floor_pct)

```

\

## Figure 6

```{r}

#==================
# Helper: overlap between 2 or 3 sets of intervals
#==================
overlap_2way <- function(df1, df2) {
  total <- 0
  if (nrow(df1) > 0 && nrow(df2) > 0) {
    for (i in seq_len(nrow(df1))) {
      for (j in seq_len(nrow(df2))) {
        start <- max(df1$xmin[i], df2$xmin[j])
        end   <- min(df1$xmax[i], df2$xmax[j])
        total <- total + max(0, end - start)
      }
    }
  }
  total
}

overlap_3way <- function(df1, df2, df3) {
  total <- 0
  if (nrow(df1) > 0 && nrow(df2) > 0 && nrow(df3) > 0) {
    for (i in seq_len(nrow(df1))) {
      for (j in seq_len(nrow(df2))) {
        for (k in seq_len(nrow(df3))) {
          start <- max(df1$xmin[i], df2$xmin[j], df3$xmin[k])
          end   <- min(df1$xmax[i], df2$xmax[j], df3$xmax[k])
          total <- total + max(0, end - start)
        }
      }
    }
  }
  total
}

#==================
# Main loop
#==================
overlap_relay <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {
  #------------------
  # Infant carrying
  #------------------
  infant <- subset(infant_caregiver, trip_id == nursing_woman)
  infant$frac <- infant$duration / sum(infant$duration)
  infant$xmax <- cumsum(infant$frac)
  infant$xmin <- c(0, head(infant$xmax, -1))
  #------------------
  # Basket carrying
  #------------------
  basket <- subset(basket_carrier, trip_id == nursing_woman)
  basket$frac <- basket$duration / sum(basket$duration)
  basket$xmax <- cumsum(basket$frac)
  basket$xmin <- c(0, head(basket$xmax, -1))
  #------------------
  # Foraging
  #------------------
  forage <- subset(foraging_bout, trip_id == nursing_woman)
  forage$frac <- forage$duration / sum(forage$duration)
  forage$xmax <- cumsum(forage$frac)
  forage$xmin <- c(0, head(forage$xmax, -1))
  #------------------
  # Resting
  #------------------
  rest <- subset(rest_bout, trip_id == nursing_woman)
  rest$frac <- rest$duration / sum(rest$duration)
  rest$xmax <- cumsum(rest$frac)
  rest$xmin <- c(0, head(rest$xmax, -1))
  #------------------
  # Walking
  #------------------
  walk <- subset(walking_bout, trip_id == nursing_woman)
  walk$frac <- walk$duration / sum(walk$duration)
  walk$xmax <- cumsum(walk$frac)
  walk$xmin <- c(0, head(walk$xmax, -1))

  #------------------
  # Reusable subsets
  #------------------
  girls               <- subset(infant, caregiver == "girl")
  mother_carry        <- subset(infant, caregiver == "mother")
  girls_basket        <- subset(basket, caregiver == "girl")
  mother_carry_basket <- subset(basket, caregiver == "mother")
  basket_floor        <- subset(basket, caregiver == "Ground")
  forage_true         <- subset(forage, forages == TRUE)
  rest_true           <- subset(rest, rest == TRUE)
  walk_true           <- subset(walk, walks == TRUE)

  #------------------
  # Total time (as fraction of trip) spent in each activity
  #------------------
  forage_total_frac <- sum(forage_true$frac)
  rest_total_frac   <- sum(rest_true$frac)
  walk_total_frac   <- sum(walk_true$frac)

  #==================
  # 1. Mother forages & carries baby & basket on ground
  #==================
  overlap_Mforages_Mcares_BasketFloor <-
    overlap_3way(forage_true, mother_carry, basket_floor)

  #==================
  # 2. Mother forages, girl carries baby, basket on ground
  #==================
  overlap_Mforages_Gcares_BasketFloor <-
    overlap_3way(forage_true, girls, basket_floor)

  #==================
  # 3. Mother rests & carries baby, basket on ground
  #==================
  overlap_Mrest_Mcares_BasketFloor <-
    overlap_3way(rest_true, mother_carry, basket_floor)

  #==================
  # 4. Mother rests, girl carries baby, basket on ground
  #==================
  overlap_Mrest_Gcares_BasketFloor <-
    overlap_3way(rest_true, girls, basket_floor)

  #==================
  # 5. Mother travels & carries baby, girl carries basket
  #==================
  overlap_Mwalk_Mcares_Gbasketcarry <-
    overlap_3way(walk_true, mother_carry, girls_basket)

  #==================
  # 6. Mother travels & carries basket, girl carries baby
  #==================
  overlap_Mwalk_MbasketCarry_Gcares <-
    overlap_3way(walk_true, mother_carry_basket, girls)

  #==================
  # 7. Mother travels & carries basket & carries baby
  #==================
  overlap_Mwalk_MbasketCarry_Mcares <-
    overlap_3way(walk_true, mother_carry_basket, mother_carry)

  #------------------
  # Normalized versions (overlap as a fraction of time spent in
  # the relevant activity, rather than fraction of the whole trip)
  # NA if the mother didn't spend any time in that activity at all
  #------------------
  norm_Mforages_Mcares_BasketFloor <-
    ifelse(forage_total_frac > 0, overlap_Mforages_Mcares_BasketFloor / forage_total_frac, NA)
  norm_Mforages_Gcares_BasketFloor <-
    ifelse(forage_total_frac > 0, overlap_Mforages_Gcares_BasketFloor / forage_total_frac, NA)
  norm_Mrest_Mcares_BasketFloor <-
    ifelse(rest_total_frac > 0, overlap_Mrest_Mcares_BasketFloor / rest_total_frac, NA)
  norm_Mrest_Gcares_BasketFloor <-
    ifelse(rest_total_frac > 0, overlap_Mrest_Gcares_BasketFloor / rest_total_frac, NA)
  norm_Mwalk_Mcares_Gbasketcarry <-
    ifelse(walk_total_frac > 0, overlap_Mwalk_Mcares_Gbasketcarry / walk_total_frac, NA)
  norm_Mwalk_MbasketCarry_Gcares <-
    ifelse(walk_total_frac > 0, overlap_Mwalk_MbasketCarry_Gcares / walk_total_frac, NA)
  norm_Mwalk_MbasketCarry_Mcares <-
    ifelse(walk_total_frac > 0, overlap_Mwalk_MbasketCarry_Mcares / walk_total_frac, NA)

  #==================
  # Save results
  #==================
  overlap_relay <- rbind(
    overlap_relay,
    data.frame(
      trip_id = nursing_woman,

      # raw fractions (of whole trip), as before
      overlap_Mforages_Mcares_BasketFloor_frac = overlap_Mforages_Mcares_BasketFloor,
      overlap_Mforages_Gcares_BasketFloor_frac = overlap_Mforages_Gcares_BasketFloor,
      overlap_Mrest_Mcares_BasketFloor_frac    = overlap_Mrest_Mcares_BasketFloor,
      overlap_Mrest_Gcares_BasketFloor_frac    = overlap_Mrest_Gcares_BasketFloor,
      overlap_Mwalk_Mcares_Gbasketcarry_frac   = overlap_Mwalk_Mcares_Gbasketcarry,
      overlap_Mwalk_MbasketCarry_Gcares_frac   = overlap_Mwalk_MbasketCarry_Gcares,
      overlap_Mwalk_MbasketCarry_Mcares_frac   = overlap_Mwalk_MbasketCarry_Mcares,

      # activity totals, for reference / sanity-checking
      forage_total_frac = forage_total_frac,
      rest_total_frac   = rest_total_frac,
      walk_total_frac   = walk_total_frac,

      # normalized versions: overlap as fraction of time in that activity
      norm_Mforages_Mcares_BasketFloor = norm_Mforages_Mcares_BasketFloor,
      norm_Mforages_Gcares_BasketFloor = norm_Mforages_Gcares_BasketFloor,
      norm_Mrest_Mcares_BasketFloor    = norm_Mrest_Mcares_BasketFloor,
      norm_Mrest_Gcares_BasketFloor    = norm_Mrest_Gcares_BasketFloor,
      norm_Mwalk_Mcares_Gbasketcarry   = norm_Mwalk_Mcares_Gbasketcarry,
      norm_Mwalk_MbasketCarry_Gcares   = norm_Mwalk_MbasketCarry_Gcares,
      norm_Mwalk_MbasketCarry_Mcares   = norm_Mwalk_MbasketCarry_Mcares
    )
  )
}

overlap_relay

overlap_relay_girls_only <- overlap_relay[trip_with_girls,]

summary(overlap_relay_girls_only$norm_Mforages_Mcares_BasketFloor)
summary(overlap_relay_girls_only$norm_Mforages_Gcares_BasketFloor)
summary(overlap_relay_girls_only$norm_Mrest_Mcares_BasketFloor)
summary(overlap_relay_girls_only$norm_Mrest_Gcares_BasketFloor)
summary(overlap_relay_girls_only$norm_Mwalk_Mcares_Gbasketcarry)
summary(overlap_relay_girls_only$norm_Mwalk_MbasketCarry_Gcares)
summary(overlap_relay_girls_only$norm_Mwalk_MbasketCarry_Mcares)

length(which(overlap_relay_girls_only$norm_Mforages_Mcares_BasketFloor > 0))
length(which(overlap_relay_girls_only$norm_Mforages_Gcares_BasketFloor > 0))
length(which(overlap_relay_girls_only$norm_Mrest_Mcares_BasketFloor > 0))
length(which(overlap_relay_girls_only$norm_Mrest_Gcares_BasketFloor > 0))
length(which(overlap_relay_girls_only$norm_Mwalk_Mcares_Gbasketcarry > 0))
length(which(overlap_relay_girls_only$norm_Mwalk_MbasketCarry_Gcares > 0))
length(which(overlap_relay_girls_only$norm_Mwalk_MbasketCarry_Mcares > 0))


```

\

## Suppl. Table 3

```{r}

#==================
# Caregiver x Activity summary table
#==================

# Get all caregiver categories actually present in the data
caregiver_levels <- sort(unique(as.character(infant_caregiver$caregiver)))

# Long-format results: one row per trip x caregiver x activity
caregiver_activity_long <- data.frame()

for (nursing_woman in focal_follow_relay_coop$ID_focal) {

  #------------------
  # Infant carrying
  #------------------
  infant <- subset(infant_caregiver, trip_id == nursing_woman)
  infant$frac <- infant$duration / sum(infant$duration)
  infant$xmax <- cumsum(infant$frac)
  infant$xmin <- c(0, head(infant$xmax, -1))

  #------------------
  # Foraging / Resting / Walking (same as before)
  #------------------
  forage <- subset(foraging_bout, trip_id == nursing_woman)
  forage$frac <- forage$duration / sum(forage$duration)
  forage$xmax <- cumsum(forage$frac)
  forage$xmin <- c(0, head(forage$xmax, -1))

  rest <- subset(rest_bout, trip_id == nursing_woman)
  rest$frac <- rest$duration / sum(rest$duration)
  rest$xmax <- cumsum(rest$frac)
  rest$xmin <- c(0, head(rest$xmax, -1))

  walk <- subset(walking_bout, trip_id == nursing_woman)
  walk$frac <- walk$duration / sum(walk$duration)
  walk$xmax <- cumsum(walk$frac)
  walk$xmin <- c(0, head(walk$xmax, -1))

  forage_true <- subset(forage, forages == 1)
  rest_true   <- subset(rest, rest == 1)
  walk_true   <- subset(walk, walks == 1)

  forage_total_frac <- sum(forage_true$frac)
  rest_total_frac   <- sum(rest_true$frac)
  walk_total_frac   <- sum(walk_true$frac)

  activities <- list(
    foraging   = list(bouts = forage_true, total = forage_total_frac),
    resting    = list(bouts = rest_true,   total = rest_total_frac),
    travelling = list(bouts = walk_true,   total = walk_total_frac)
  )

  #------------------
  # Loop over caregiver categories x activities
  #------------------
  for (cg in caregiver_levels) {
    cg_bouts <- subset(infant, caregiver == cg)

    for (act_name in names(activities)) {
      act_bouts <- activities[[act_name]]$bouts
      act_total <- activities[[act_name]]$total

      ov <- overlap_2way(act_bouts, cg_bouts)
      norm_val <- ifelse(act_total > 0, ov / act_total, NA)

      caregiver_activity_long <- rbind(
        caregiver_activity_long,
        data.frame(
          trip_id   = nursing_woman,
          caregiver = cg,
          activity  = act_name,
          norm_frac = norm_val
        )
      )
    }
  }
}

#==================
# Average across trips -> wide table (caregiver rows x activity columns)
#==================
caregiver_activity_summary <- aggregate(
  norm_frac ~ caregiver + activity,
  data = caregiver_activity_long,
  FUN = mean, na.rm = TRUE
)

caregiver_activity_table <- reshape(
  caregiver_activity_summary,
  idvar = "caregiver",
  timevar = "activity",
  direction = "wide"
)

# clean up column names: norm_frac.foraging -> foraging, etc.
names(caregiver_activity_table) <- gsub("norm_frac\\.", "", names(caregiver_activity_table))

# reorder columns nicely
caregiver_activity_table <- caregiver_activity_table[, c("caregiver", "foraging", "resting", "travelling")]

# convert to percentages, rounded
caregiver_activity_table$foraging   <- round(caregiver_activity_table$foraging   * 100, 1)
caregiver_activity_table$resting    <- round(caregiver_activity_table$resting    * 100, 1)
caregiver_activity_table$travelling <- round(caregiver_activity_table$travelling * 100, 1)

caregiver_activity_table

```

###### Table 1: Assistance during foraging trips

This analysis calculates the frequency and duration of infant-care and
basket-carrying assistance during focal women's foraging trips.

For age- and gender-specific categories, calculations include only trips in
which at least one member of that category was present. A category that was
present but did not provide assistance contributes a value of zero. Trips in
which the category was absent are excluded.

For the pooled ``Any other group member'' category, calculations are restricted
to the 17 trips in which the focal woman was accompanied by at least one other
group member.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
```

## 1. Prepare infant-care data

The unknown infant-care category (`?`) is combined with `Ground`.

```{r prepare-infant-data}
infant_caregiver2 <- infant_caregiver %>%
  mutate(
    caregiver = ifelse(caregiver == "?", "Ground", caregiver)
  )
```

## 2. Identify trips in which each category was present

```{r category-presence}
trips_women_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_women > 0
]

trips_men_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_men > 0
]

trips_girls_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_girls > 0
]

trips_boys_present <- focal_follow_relay_coop$ID_focal[
  focal_follow_relay_coop$n_boys > 0
]

trips_accompanied <- focal_follow_relay_coop %>%
  filter(
    n_women > 0 |
    n_men > 0 |
    n_girls > 0 |
    n_boys > 0
  ) %>%
  pull(ID_focal)
```

### Check denominators

These should reproduce the denominators used in Table 1.

```{r check-denominators}
length(trips_accompanied)    # expected: 17
length(trips_women_present)  # expected: 14
length(trips_men_present)    # expected: 6
length(trips_girls_present)  # expected: 13
length(trips_boys_present)   # expected: 11
```

# Infant care

## 3. Calculate trip-level infant-care percentages

For each trip, calculate the percentage of total trip duration associated with
each infant-care category.

```{r infant-trip-percentages}
caregiver_trip <- infant_caregiver2 %>%
  group_by(trip_id, caregiver) %>%
  summarise(
    duration = sum(duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(trip_id) %>%
  mutate(
    percent = duration / sum(duration, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```

## 4. Infant care by category, conditional on presence

If a category was present but did not provide infant care, the trip contributes
0%. If the category was absent, the trip is excluded.

```{r infant-summary-function}
summarise_care_when_present <- function(category, trips_present) {

  x <- data.frame(trip_id = trips_present) %>%
    left_join(
      caregiver_trip %>%
        filter(caregiver == category) %>%
        select(trip_id, percent),
      by = "trip_id"
    ) %>%
    mutate(
      percent = replace_na(percent, 0)
    )

  x %>%
    summarise(
      caregiver = category,
      N_present = n(),
      N_helped = sum(percent > 0),
      pct_trips_helped = N_helped / N_present * 100,
      mean = mean(percent),
      sd = sd(percent),
      min = min(percent),
      max = max(percent)
    )
}
```

```{r infant-category-summary}
caregiver_summary_present <- bind_rows(
  summarise_care_when_present("woman",  trips_women_present),
  summarise_care_when_present("father", trips_men_present),
  summarise_care_when_present("girl",   trips_girls_present),
  summarise_care_when_present("boy",    trips_boys_present)
)

caregiver_summary_present
```

Expected frequencies:

- Other women: 8/14
- Fathers: 4/6
- Girls: 9/13
- Boys: 4/11

## 5. Infant-care assistance from any other group member

`mother` and `Ground` are excluded when defining assistance from another
individual.

```{r infant-any-helper}
infant_help_trip <- infant_caregiver2 %>%
  filter(trip_id %in% trips_accompanied) %>%
  mutate(
    other_help = !caregiver %in% c("mother", "Ground")
  ) %>%
  group_by(trip_id) %>%
  summarise(
    total_duration = sum(duration, na.rm = TRUE),
    other_duration = sum(duration[other_help], na.rm = TRUE),
    percent_other = other_duration / total_duration * 100,
    .groups = "drop"
  )

infant_help_summary <- infant_help_trip %>%
  summarise(
    N = n(),
    N_helped = sum(other_duration > 0),
    percent_trips_helped = N_helped / N * 100,
    mean = mean(percent_other),
    sd = sd(percent_other),
    min = min(percent_other),
    max = max(percent_other)
  )

infant_help_summary
```

Expected result: **17/17 (100%)**, mean = **38.9%**, SD = **27.0%**,
range = **1.9--100.0%**.

# Basket carrying

## 6. Calculate trip-level basket-carrying percentages

```{r basket-trip-percentages}
basket_trip <- basket_carrier %>%
  group_by(trip_id, caregiver) %>%
  summarise(
    duration = sum(duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(trip_id) %>%
  mutate(
    percent = duration / sum(duration, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```

## 7. Basket carrying by category, conditional on presence

```{r basket-summary-function}
summarise_basket_when_present <- function(category, trips_present) {

  x <- data.frame(trip_id = trips_present) %>%
    left_join(
      basket_trip %>%
        filter(caregiver == category) %>%
        select(trip_id, percent),
      by = "trip_id"
    ) %>%
    mutate(
      percent = replace_na(percent, 0)
    )

  x %>%
    summarise(
      caregiver = category,
      N_present = n(),
      N_carried = sum(percent > 0),
      pct_trips_carried = N_carried / N_present * 100,
      mean = mean(percent),
      sd = sd(percent),
      min = min(percent),
      max = max(percent)
    )
}
```

```{r basket-category-summary}
basket_summary_present <- bind_rows(
  summarise_basket_when_present("woman",  trips_women_present),
  summarise_basket_when_present("father", trips_men_present),
  summarise_basket_when_present("girl",   trips_girls_present),
  summarise_basket_when_present("boy",    trips_boys_present)
)

basket_summary_present
```

Expected frequencies:

- Other women: 4/14
- Fathers: 0/6
- Girls: 8/13
- Boys: 3/11

## 8. Basket-carrying assistance from any other group member

```{r basket-any-helper}
basket_help_trip <- basket_carrier %>%
  filter(trip_id %in% trips_accompanied) %>%
  mutate(
    other_help = !caregiver %in% c("mother", "Ground")
  ) %>%
  group_by(trip_id) %>%
  summarise(
    total_duration = sum(duration, na.rm = TRUE),
    other_duration = sum(duration[other_help], na.rm = TRUE),
    percent_other = other_duration / total_duration * 100,
    .groups = "drop"
  )

basket_help_summary <- basket_help_trip %>%
  summarise(
    N = n(),
    N_helped = sum(other_duration > 0),
    percent_trips_helped = N_helped / N * 100,
    mean = mean(percent_other),
    sd = sd(percent_other),
    min = min(percent_other),
    max = max(percent_other)
  )

basket_help_summary
```

Expected result: **12/17 (70.6%)**, mean = **18.5%**, SD = **21.2%**,
range = **0.0--67.7%**.

# Final checks

The frequencies calculated from the duration data should reproduce the
frequencies reported in Table 1.

```{r final-checks}
caregiver_summary_present %>%
  select(caregiver, N_present, N_helped, pct_trips_helped)

basket_summary_present %>%
  select(caregiver, N_present, N_carried, pct_trips_carried)

infant_help_summary

basket_help_summary
```

